---
title: "TERRE MatrixEQTL Preparation"
output: html_notebook
---

The manuscript "A Parkinson’s disease genetic risk score associates with blood DNAm on chromosome 17" explores how a genetic risk score (GRS) for Parkinson's disease (PD) associates with blood DNA methylation in the TERRE study of French agricultural workers, and evaluates GRS associations with DNAm in each sex as well as their sensitivity to non-genetic factors. This was originally spearheaded by Will Casazza and was posted as a pre-print on medRxiv in 2023. The current revision (2024, led by Sam Schaffner) aims to address previous reviewer feedback and prepare the manuscript for re-submission.


This code prepares genetic, methylation, and meta data for input into MatrixEQTL to calculate mQTLs in TERRE.


Last updated July 28, 2024 by Samantha Schaffner

Based on code by Will Casazza

```{r setup, include=FALSE, eval=F}
library(tidyverse)
library(data.table)
library(minfi)
```

# Read in TERRE data
DNA methylation (updated to 2022-processed), genotype, and meta data (with 2022 cell type PCs).
```{r terre meta data, eval=F}
terre_ancestry <- fread("~/SHARE_DECIPHER/pd_grs_ewas_integration/raw_data.geno.maf.mind.sex_check.het_filter.ibd_filter.eigenvec")

terre_meta <- fread("/home1/NEURO/SHARE_DECIPHER/terre_meta_master.csv")
#terre_ctp <- fread("/home1/NEURO/SHARE_DECIPHER/processed_DNAm_data/2021/TERRE_processed_2021/meta_TERRE_funnorm_robCTP.txt")
terre_ctp <- read.csv("~/PD_GRS/meta_for_GRS_EWAS.csv") # with cell type PCs from 2022 DNAm data

terre_meta$IID <- gsub(".*_(PAE.*)", "\\1", terre_meta$IID)
terre_meta <- terre_meta %>%
  right_join(terre_ancestry, by = c("IID" = "V2")) %>%
  left_join(terre_ctp[, c("patient", "CTP_PC1", "CTP_PC2", "CTP_PC3", "CTP_PC4", "CTP_PC5", "CTP_PC6")], by = "patient")
colnames(terre_meta) #contains cell type and genotype PCs, FID, IID, and other meta variables
#245 observations

# Loading in DNAm data
# Update from previous script that used 2021-processed data
load("~/SHARE_DECIPHER/processed_DNAm_data/2022/TERRE_processed_2022/1-TERRE_RG_filtered.RData")
betas <- getBeta(PD_RG_filtered)
rm(PD_RG_filtered)

genotype <- fread("~/SHARE_DECIPHER/pd_grs_ewas_integration/TERRE_QC/raw_data.imputed.r2_30.maf_mismatch.traw")
```

## Match Samples
```{r match samples, eval=F}
colnames(genotype) <- gsub(".*_(PAE.*)", "\\1", colnames(genotype))
colnames(genotype)
setcolorder(genotype, neworder = c(colnames(genotype)[1:6], terre_meta$IID))
dim(betas <- betas[, colnames(betas) %in% terre_meta$patient]) #803777    219

# new - match terre meta to betas (219 betas, 245 terre meta)
dim(terre_meta <- terre_meta[match(colnames(betas),terre_meta$patient),]) #219  96

# rename betas sample names with IID
colnames(betas) <- terre_meta$IID[match(colnames(betas), terre_meta$patient)]
all(colnames(betas)==terre_meta$IID) #TRUE

# match genotype and betas data
geno_df <- data.frame(genotype)
ncol(geno_df <- cbind(geno_df[,1:6], geno_df[,match(colnames(betas),colnames(geno_df))])) #225
all(colnames(betas) == colnames(geno_df)[-c(1:6)]) #TRUE
```

## Compute DNAm PCs
```{r dnam pcs, eval=F}
methy_PC <- prcomp(t(betas), center = T, rank. = 20)
```

## Write Covariates

Covariates used for mQTL analysis typically include genotype PCs, sex, methylation PCs, and sometimes methylation batch and cell type variables. Example references:

![GODMC](https://github.com/MRCIEU/godmc/wiki/Phenotype-and-covariate-data): Sex, age, DNAm chip (and other batch variables), height, BMI, cell counts, genotype PCs (created by pipeline), methylation PCs (created by pipeline)

![Oliva et al., Nat Genet 2023](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10249665/): Sex, 5 genotype PCs, 2 genotype batch variables

![Li et al., Clin Epigenetics 2022](https://clinicalepigeneticsjournal.biomedcentral.com/articles/10.1186/s13148-022-01385-6): Sex, 5 genotype PCs, DNAm batch variables

![Ohlei et al., medRxiv 2023](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10775408/): Sex, 10 genotype PCs, 5 DNAm PCs, sample processing date


The code previously written by Will here included 3 genotype PCs, age, sex, head trauma, cell type PCs, and PD status. There was also an analysis with 10 DNAm PCs.

Here I will include the same covariates minus head trauma and DNAm PCs and plus DNAm plate and row variables. This will match the variables adjusted for in the updated EWAS. Not sure why head trauma was previously included here - check with Will.


```{r write covars, eval=F}
# data need to be in numeric format

terre_covar <- terre_meta[,c("V3", "V4", "V5", "PD", "age", "men", "CTP_PC1", "CTP_PC2", "CTP_PC3", "CTP_PC4", "CTP_PC5","CTP_PC6","SentrixPosition_A","plate")]
str(terre_covar)
terre_covar$SentrixPosition_A <- as.integer(as.factor(terre_covar$SentrixPosition_A))
summary(as.factor(terre_covar$SentrixPosition_A))
# 1  2  3  4  5  6  7  8 
#33 31 30 22 22 34 27 20 

write_delim(t(cbind(terre_covar)) %>% as.data.frame() %>% rownames_to_column("id"), "~/PD_GRS/mQTL/covariates_CTP_PD.txt")

write_delim(t(cbind(terre_covar[terre_covar$men==1,-("men")])) %>% as.data.frame() %>% rownames_to_column("id"), "~/PD_GRS/mQTL/male_covariates_CTP_PD.txt")

write_delim(t(cbind(terre_covar[terre_covar$men==0,-("men")])) %>% as.data.frame() %>% rownames_to_column("id"), "~/PD_GRS/mQTL/female_covariates_CTP_PD.txt")
```

## Write genotype and DNAm data
```{r write geno dnam, eval=F}
#methy_annot <- fread("~/MethylationEPIC_v-1-0_B4.csv", skip = 7)
load("~/kobor_shared_coding_resource/EPIC_fdat.RData")

# SNP POS
geno_df$CHR <- paste0("chr", geno_df$CHR)
write_delim(geno_df[, c("SNP", "CHR", "POS")], "~/PD_GRS/mQTL/snp_pos.txt")

# SNPs
geno_vars <- colnames(geno_df)[-c(1, 3:6)] #220 (219)
geno_vars_m <- geno_vars[-(1 + which(terre_meta$men == 0))] #120 (119)
geno_vars_f <- geno_vars[-(1 + which(terre_meta$men == 1))] #101 (100)
fwrite(geno_df[, geno_vars], "~/PD_GRS/mQTL/all_imputed_matrixeQTL.txt", sep = " ", quote = F)
fwrite(geno_df[, geno_vars_f], "~/PD_GRS/mQTL/female_all_imputed_matrixeQTL.txt", sep = " ", quote = F)
fwrite(geno_df[, geno_vars_m], "~/PD_GRS/mQTL/male_all_imputed_matrixeQTL.txt", sep = " ", quote = F)

# Methy POS
pos_dat <- data.frame(geneid=fData_EPIC$TargetID, chr=paste0("chr", fData_EPIC$CHR), s1=fData_EPIC$MAPINFO, s2=fData_EPIC$MAPINFO)
dim(pos_dat <- pos_dat[match(rownames(betas), pos_dat$geneid),]) #803777      5
fwrite(pos_dat, "~/PD_GRS/mQTL/probe_pos.txt", sep = " ", quote = F)

# methy
#betas_write <- betas %>% as.data.frame() %>% rownames_to_column("cpg")

fwrite(betas %>% as.data.frame() %>% rownames_to_column("cpg"), "~/PD_GRS/mQTL/methylation.txt", sep = " ", quote = F)
fwrite(betas[, terre_meta$men == 1] %>% as.data.frame() %>% rownames_to_column("cpg"), "~/PD_GRS/mQTL/male_methylation.txt", sep = " ", quote = F)
fwrite(betas[, terre_meta$men == 0] %>% as.data.frame() %>% rownames_to_column("cpg"), "~/PD_GRS/mQTL/female_methylation.txt", sep = " ", quote = F)

# SNP POS
write_delim(geno_df[geno_df$CHR == "chr21", c("SNP", "CHR", "POS")], "~/PD_GRS/mQTL/snp_pos_chr21.txt")

# SNPs
geno_vars <- colnames(geno_df)[-c(1, 3:6)]
write_delim(geno_df[geno_df$chr=="CHR21", geno_vars], "~/PD_GRS/mQTL/all_imputed_matrixeQTL_chr21.txt")

# Methy POS
write_delim(pos_dat[pos_dat$chr == "chr21",], "~/PD_GRS/mQTL/probe_pos_chr21.txt")
length(chr21_cpg <- pos_dat[pos_dat$chr == "chr21","Name"]) #9630

# methy
write_delim(betas %>% as.data.frame() %>% rownames_to_column("cpg") %>% filter(cpg %in% chr21_cpg), "~/PD_GRS/mQTL/methylation_chr21.txt")
```

All inputs for MatrixEQTL are now generated. Refer to scripts "run_matrixEQTL.R" and "run_mQTL_terre.sh" to generate results prior to the next section.

# Checking number of PCs
```{r pc check, eval=F}
pc_res <- lapply(0:20, function(i) fread(sprintf("terre_data/cis_all_impute_mQTL_results_%d_methy_PC_chr21.txt.gz", i)))
ctp_res <- fread("terre_data/cis_all_impute_mQTL_results_CTP_chr21.txt.gz")
results_df <- data.frame(PC = 0:20, nhit = sapply(pc_res, function(dt) nrow(dt[FDR < 0.05])))
ggplot(results_df, aes(PC, nhit)) +
  geom_point() +
  geom_smooth() +
  geom_hline(yintercept = nrow(ctp_res[FDR < 0.05]), linetype = "dashed") +
  annotate(geom = "text", y = 240500, x = 10, label = "Hits detected with CTP PC") +
  ggtitle("mQTLs detected on chr21 Vs Methylation PCs included")

res <- fread("~/PD_GRS/mQTL/cis_all_impute_mQTL_results.txt.gz")

results_df <- data.frame(PC = 0:20, nhit = sapply(pc_res, function(dt) nrow(dt[`p-value` < (0.05 / .N), .(unique(gene))])))
a1 <- ggplot(results_df, aes(PC, nhit)) +
  geom_point() +
  geom_smooth() +
  geom_hline(yintercept = nrow(ctp_res[FDR < 0.05, .(unique(gene))]), linetype = "dashed") +
  annotate(geom = "text", y = 4200, x = 10, label = "Hits detected with cell type 5 proportion PCs") +
  theme_classic() +
  labs(y = "# mQTLs Detected", x = "# PCs Included") #+ ggtitle("CpG sites on chr21 with one mQTL Vs Methylation PCs included")
```
