---
title: "PRS Analysis"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(data.table)
library(limma)
library(minfi)
library(rcompanion)
library(readxl)
knitr::opts_chunk$set(echo = TRUE)
```
# PD risk score data and matching
```{r}
risk_scores <- read_xlsx("~/pd_risk_score/Supplementary materials/Table S2. Detailed summary statistics on all nominated risk variants, known and novel_.xlsx") %>% mutate(CHR = as.numeric(CHR), COUNTED = toupper(`Effect allele`), ALT = toupper(`Other allele`), POS = BP) %>% filter(!is.na(CHR))
risk_scores %>% select(SNP,COUNTED, `Beta, all studies`) %>%
  write_delim("pd_prs_file.txt", col_names = FALSE)

```

# TERRE PRS Analysis
Generate PRS using `plink2 --score`:
```{bash}
~/plink2 --bfile ~/genotype_qc/TERRE_QC/raw_data.imputed.r2_30.hard_call.maf_mismatch\
  --score pd_prs_file.txt\
  --out terre_prs
```
## Loading in and preparing data
```{r}
terre_prs <- fread("terre_prs.sscore")
terre_pcs <- fread("~/genotype_qc/TERRE_QC/raw_data.geno.maf.mind.sex_check.het_filter.ibd_filter.eigenvec")
terre_metadata <- read.csv("/home1/NEURO/SHARE_DECIPHER/terre_meta_master.csv") %>% 
  mutate(
    FID = gsub("(PAE_[0-9]*_[1-9]*)_.*","\\1",FID),
    IID = gsub("(PAE_[0-9]*_[1-9]*)_.*","\\1",IID)
  )
all_data <- terre_prs %>% left_join(terre_metadata, by=c("IID")) %>% left_join(terre_pcs,by=c("IID"= "V1"))
```
## Association analysis
```{r}
pd_h2 <- function(R2O,case_prop,prevalence=0.22){
  K <- prevalence
  P <- case_prop
  thd <- -1 * qnorm(K,0,1)
  zv <- dnorm(thd) #z (normal density)
  mv <- zv/K #mean liability for case
  theta <- mv*(P-K)/(1-K)*(mv*(P-K)/(1-K)-thd) #θ in equation 
  cv <- K*(1-K)/zv^2*K*(1-K)/(P*(1-P)) #C in
  return(R2O*cv/(1+R2O*theta*cv))
}
fit1 <- glm(PD ~ SCORE1_AVG + men + age + V3 + V4 + V5, data=all_data,family="binomial")
summary(fit1)
fit1_null <- glm(PD ~  men + age + V3 + V4 + V5, data=all_data,family="binomial")
summary(fit1_null)
nagelkerke(fit1,null=fit1_null)
#liability R2
lin1 <- lm(PD ~ SCORE1_AVG + men + age + V3 + V4 + V5, data=all_data)
lin2 <- lm(PD ~ men + age + V3 + V4 + V5, data=all_data)
R2 <- (deviance(lin2) - deviance(lin1)) / deviance(lin2)
h2 <- pd_h2(R2,sum(all_data$status)/nrow(all_data),0.01)
fit2 <- glm(PD ~ SCORE1_AVG*men + V3+V4+V5, data = all_data, family="binomial")
summary(fit2)
fit3 <- glm(PD~SCORE1_AVG*smoking + V3+V4+V5, data=all_data, family="binomial")
summary(fit3)
fit4 <- glm(PD~SCORE1_AVG*pesticides + V3+V4+V5, data=all_data, family="binomial")
summary(fit4)

fit_null <- glm(PD ~ SCORE1_AVG, data=all_data, family="binomial")
summary(fit_null)
```

```{r}
ggplot(all_data %>% mutate(PD= ifelse(PD,"CASE","CONTROL")),aes(as.factor(PD),SCORE1_AVG)) +geom_boxplot()+ geom_jitter(height=0, width=0.1)+ labs(y="Polygenic Risk Score", x = "Parkinson's Diagnosis") + ggtitle(bquote('Nagelkerke Pseudo-'*R^2*': 0.0611' ))
ggplot(all_data,aes(factor(men),SCORE1_AVG)) +geom_boxplot()+ geom_jitter(height=0, width=0.1) 
t.test(all_data$SCORE1_AVG[as.logical(all_data$PD)],all_data$SCORE1_AVG[!as.logical(all_data$PD)],alternative = "greater")
t.test(all_data$SCORE1_AVG[as.logical(all_data$men)],all_data$SCORE1_AVG[!as.logical(all_data$men)])
```

## PRS vs DNAm

```{r}
# load("/home1/NEURO/SHARE_DECIPHER/processed_DNAm_data/TERRE/TERRE_betas_CTC.RData")# terre_betas_CTC
load("/home1/NEURO/SHARE_DECIPHER/processed_DNAm_data/TERRE/TERRE_betas_combat.RData")# betas_sub
# load("/home1/NEURO/SHARE_DECIPHER/processed_DNAm_data/TERRE/TERRE_funnorm_processed/betas_combat.RData")
# betas_sub<- betas_combat
```
Let's check how the data looks for a random subject:
```{r}
ggplot(betas_sub[,c(1,2,3,4,5)] %>% as.data.table(keep.rownames = T) %>% melt(id.vars="rn",value.name = "betas", variable.name="subject"),aes(betas,color=subject)) + geom_density()
```

### Match DNA, PRS, and metadata
```{r}
reg_all_data <- all_data[na.omit(match(colnames(betas_sub),all_data$patient)),]
reg_all_data$sex <- ifelse(reg_all_data$men,"Male","Female")
reg_all_data$PD <- ifelse(reg_all_data$PD,"CASE","CONTROL")
betas_sub <- betas_sub[,colnames(betas_sub) %in% reg_all_data$patient]
all(colnames(betas_sub) == reg_all_data$patient)
```
### TERRE PRS regression vs methylation

#### Differential methylation with PRS vs PD
```{r}
na_row <- !apply(reg_all_data[,c("SCORE1_AVG","age","plate","sex")],1,function(x)any(is.na(x)))
design_prs <- model.matrix(~1+SCORE1_AVG+age+plate+sex+V3+V4+V5, data=reg_all_data[na_row,])
prs_fit <- lmFit(betas_sub[,na_row],design_prs)
prs_fit <- eBayes(prs_fit)
# Cases only
case_select <-na_row & reg_all_data$PD =="CASE"
design_prs_case <- model.matrix(~1+SCORE1_AVG+age+plate+sex+V3+V4+V5, data=reg_all_data[case_select,])
prs_fit_case <- lmFit(betas_sub[,case_select],design_prs_case)
prs_fit_case <- eBayes(prs_fit_case)
# Controls only
control_select <-na_row & reg_all_data$PD =="CONTROL"
design_prs_control <- model.matrix(~1+SCORE1_AVG+age+plate+sex+V3+V4+V5, data=reg_all_data[control_select,])
prs_fit_control <- lmFit(betas_sub[,control_select],design_prs_control)
prs_fit_control <- eBayes(prs_fit_control)

design_pd <- model.matrix(~1+PD+age+plate+sex, data=reg_all_data[na_row,])
pd_fit <- lmFit(betas_sub[,na_row],design_pd)
pd_fit <- eBayes(pd_fit)
pd_res <- topTable(pd_fit,coef=2,number = Inf)
prs_res <- topTable(prs_fit,coef=2,number = Inf)
control_res <- topTable(prs_fit_control,coef=2,number = Inf)
case_res <- topTable(prs_fit_case,coef=2,number = Inf)
```


#### Permutation test for checking if reason for discrepancy between cases and controls is sample sized
```{r}
library(parallel)
num_control <- sum(control_select)
num_cases <- sum(case_select)

run_at_sample_size <- function(size){
  to_select <- sample((1:length(na_row))[na_row],size)
  design_prs <- model.matrix(~1+SCORE1_AVG+age+plate+sex+V3+V4+V5, data=reg_all_data[to_select,])
  prs_fit <- lmFit(betas_sub[,to_select],design_prs)
  prs_fit <- eBayes(prs_fit)
  topTable(prs_fit,coef=2,number = Inf)
}


run_at_sample_size_control <- function(size){
  to_select <- sample((1:length(na_row))[control_select],size)
  design_prs <- model.matrix(~1+SCORE1_AVG+age+plate+sex+V3+V4+V5, data=reg_all_data[to_select,])
  prs_fit <- lmFit(betas_sub[,to_select],design_prs)
  prs_fit <- eBayes(prs_fit)
  topTable(prs_fit,coef=2,number = Inf)
}

res_control_size <- lapply(1:10,function(x) run_at_sample_size(num_control))
res_case_size <- lapply(1:10,function(x) run_at_sample_size(num_cases))

res_case_size_control <- lapply(1:10,function(x) run_at_sample_size_control(num_cases))
```

```{r}
control_hits <- unlist(lapply(res_control_size,function(x) nrow(subset(x,adj.P.Val < 0.05 & abs(logFC) > 0.03))))

case_hits <- unlist(lapply(res_case_size,function(x) nrow(subset(x,adj.P.Val < 0.05 & abs(logFC) > 0.03))))
case_size_control_hits <-unlist(lapply(res_case_size_control,function(x) nrow(subset(x,adj.P.Val < 0.05 & abs(logFC) > 0.03))))
df <- data.frame(hits = c(mean(case_hits),mean(control_hits)), sample_size=c("Sampled to |Cases|","Sampled to |Controls|"),dev=c(sd(case_hits),sd(control_hits)))
ggplot(df,aes(x=sample_size,y=hits,ymin=hits-dev,ymax=hits+dev)) + geom_bar(stat="identity") + geom_errorbar(width=0.25) + theme_minimal() + labs(x="",y="Number of hits")
t.test(case_hits,control_hits)
```


```{r}
library(ggrepel)
annot <- fread("~/MethylationEPIC_v-1-0_B4.csv",skip = 7)
pd_annot <- pd_res %>%rownames_to_column("probe") %>% left_join(annot,by=c("probe"="Name"))%>% mutate(gene = gsub(";.*","",UCSC_RefGene_Name)) %>% mutate(gene= ifelse(gene == "",probe,gene))

prs_annot <- prs_res %>%rownames_to_column("probe") %>% left_join(annot,by=c("probe"="Name"))%>% mutate(gene = gsub(";.*","",UCSC_RefGene_Name)) %>% mutate(gene= ifelse(gene == "",probe,gene))

ggplot(prs_annot,aes(logFC,-log10(P.Value)))+ 
  geom_point() + 
  geom_text_repel(
    data=prs_annot%>% filter(abs(logFC) > 0.03 & adj.P.Val < 0.05),
    color="dodgerblue",
    mapping=aes(logFC,-log10(P.Value),label=gene)
  ) +
  geom_point(
    data=subset(prs_annot,adj.P.Val < 0.05 & abs(logFC) > 0.03),
    color = "green",
    mapping=aes(logFC,-log10(P.Value))
  ) +
  labs(x = quote(Delta~beta~Methylation)) +
  theme_minimal() +
  geom_hline(
    linetype="dashed",
    yintercept = min(-log10(prs_annot$P.Value[prs_annot$adj.P.Val <0.05]))
  ) +
  geom_vline(linetype="dashed",xintercept = 0.03) +
  geom_vline(linetype="dashed",xintercept = -0.03)+ 
  ggtitle("Cases and Controls")
# prs_annot[prs_annot$adj.P.Val <0.05,c("gene","CHR","MAPINFO","logFC","P.Value","UCSC_RefGene_Name")] %>% mutate(UCSC_RefGene_Name = sapply(UCSC_RefGene_Name,function(x)split(x)[1]))
prs_highlight <- subset(prs_annot,adj.P.Val < 0.05 & abs(logFC) > 0.03)
case_annot <- case_res %>%rownames_to_column("probe") %>% left_join(annot,by=c("probe"="Name"))%>% mutate(gene = gsub(";.*","",UCSC_RefGene_Name)) %>% mutate(gene= ifelse(gene == "",probe,gene))
ggplot(case_annot,aes(logFC,-log10(P.Value)))+
  geom_point() +
  geom_text_repel(
    data=subset(case_annot,adj.P.Val < 0.05& abs(logFC) > 0.03 & !probe %in% prs_highlight$probe),
    color="dodgerblue",
    mapping=aes(logFC,-log10(P.Value),label=gene)
  ) +
  geom_point(
    data=subset(case_annot,adj.P.Val < 0.05 & abs(logFC) > 0.03 & !probe %in% prs_highlight$probe), 
    color = "red",
    mapping=aes(logFC,-log10(P.Value))
  ) +
    geom_text_repel(
    data=case_annot %>% filter(probe %in% prs_highlight$probe),#subset(case_annot,adj.P.Val < 0.05& abs(logFC) > 0.03),
    color="dodgerblue",
    mapping=aes(logFC,-log10(P.Value),label=gene)
  ) +
  geom_point(
    data=case_annot %>% filter(probe %in% prs_highlight$probe),#subset(case_annot,adj.P.Val < 0.05 & abs(logFC) > 0.03), 
    color = "green",
    mapping=aes(logFC,-log10(P.Value))
  ) +
  labs(x = quote(Delta~beta~Methylation)) + 
  theme_minimal() + 
  geom_hline(
    linetype="dashed",
    yintercept = min(-log10(case_annot$P.Value[case_annot$adj.P.Val <0.05]))
  ) +
  geom_vline(linetype="dashed",xintercept = 0.03) +
  geom_vline(linetype="dashed",xintercept = -0.03) + 
  ggtitle("Cases")
# case_annot[case_annot$adj.P.Val <0.05,c("gene","CHR","MAPINFO","logFC","P.Value","UCSC_RefGene_Name")] %>% mutate(UCSC_RefGene_Name = sapply(UCSC_RefGene_Name,function(x)split(x)[1]))

control_annot <- control_res %>%rownames_to_column("probe") %>% left_join(annot,by=c("probe"="Name"))%>% mutate(gene = gsub(";.*","",UCSC_RefGene_Name)) %>% mutate(gene= ifelse(gene == "",probe,gene))
ggplot(control_annot,aes(logFC,-log10(P.Value)))+ 
  geom_point() +
  geom_text_repel(
    data=subset(control_annot,adj.P.Val < 0.05& abs(logFC) > 0.03 & !probe %in% prs_highlight$probe),
    color="dodgerblue",
    mapping=aes(logFC,-log10(P.Value),label=gene)
  ) +
  geom_point(
    data=subset(control_annot,adj.P.Val < 0.05 & abs(logFC) > 0.03 & !probe %in% prs_highlight$probe),
    color = "red",
    mapping=aes(logFC,-log10(P.Value))
  ) +
    geom_text_repel(
    data=control_annot %>% filter(probe %in% prs_highlight$probe),#subset(control_annot,adj.P.Val < 0.05& abs(logFC) > 0.03),
    color="dodgerblue",
    mapping=aes(logFC,-log10(P.Value),label=gene)
  ) +
  geom_point(
    data=control_annot %>% filter(probe %in% prs_highlight$probe),#subset(control_annot,adj.P.Val < 0.05 & abs(logFC) > 0.03),
    color = "green",
    mapping=aes(logFC,-log10(P.Value))
  ) +
  labs(x = quote(Delta~beta~Methylation)) + 
  theme_minimal() + geom_hline(linetype="dashed",yintercept = min(-log10(control_annot$P.Value[control_annot$adj.P.Val <0.05]))) +
  geom_vline(linetype="dashed",xintercept = 0.03) +
  geom_vline(linetype="dashed",xintercept = -0.03)+ 
  ggtitle("Controls")

# control_annot[control_annot$adj.P.Val <0.05,c("gene","CHR","MAPINFO","logFC","P.Value","UCSC_RefGene_Name")] %>% mutate(UCSC_RefGene_Name = sapply(UCSC_RefGene_Name,function(x)split(x)[1]))


ggplot(pd_annot,aes(logFC,-log10(P.Value)))+ 
  geom_point() + 
    geom_text_repel(
    data=pd_annot %>% filter(probe %in% prs_highlight$probe),#subset(pd_annot,adj.P.Val < 0.05& abs(logFC) > 0.03),
    color="dodgerblue",
    mapping=aes(logFC,-log10(P.Value),label=gene)
  ) +
  geom_point() + 
  geom_text_repel(
    data=pd_annot %>%filter(abs(logFC) > 0.03 & adj.P.Val < 0.05) %>% top_n(25,adj.P.Val),#subset(pd_annot,adj.P.Val < 0.05& abs(logFC) > 0.03),
    color="dodgerblue",
    mapping=aes(logFC,-log10(P.Value),label=gene)
  ) +
  geom_point(
    data=pd_annot %>% filter(probe %in% prs_highlight$probe),#subset(pd_annot,adj.P.Val < 0.05 & abs(logFC) > 0.03),
    color = "green",
    mapping=aes(logFC,-log10(P.Value))
  ) +
  geom_point(
    data=subset(pd_annot,adj.P.Val < 0.05 & abs(logFC) > 0.03 & !probe %in% prs_highlight$probe),
    color = "red",
    mapping=aes(logFC,-log10(P.Value))
  ) +
  labs(x = quote(Delta~beta~Methylation)) +
  theme_minimal() +
  geom_hline(
    linetype="dashed",
    yintercept = min(-log10(pd_annot$P.Value[pd_annot$adj.P.Val <0.05]))
  ) +
  geom_vline(linetype="dashed",xintercept = 0.03) +
  geom_vline(linetype="dashed",xintercept = -0.03)+ 
  ggtitle("PD Diagnosis")
```

```{r}
library(qqman)
qqman::qq(pd_res$P.Value)
manhattan(prs_annot %>% mutate(P=P.Value,SNP=gene, CHR=as.numeric(recode(CHR,X="23",Y="24",MT="26")),BP=MAPINFO),chrlabs = c(1:22,"X","Y"),annotateTop = F,annotatePval = 5e-8)
qqman::qq(prs_res$P.Value)
hist(prs_res$P.Value,breaks = 100)
hist(pd_res$P.Value,breaks=100)
```
#### What are these PRS hits?


```{r}
prs_res %>% rownames_to_column("probe") %>% filter(P.Value < (0.05/ n())) %>% left_join(annot %>% select(Name,CHR,UCSC_RefGene_Name,Relation_to_UCSC_CpG_Island,MAPINFO), by=c("probe"="Name"))%>% mutate(gene = gsub(";.*","",UCSC_RefGene_Name)) %>% mutate(gene= ifelse(gene == "",probe,gene))
pd_res %>% rownames_to_column("probe") %>% filter(P.Value < (0.05/ n()))#, abs(logFC) > 0.1)
tmp <- prs_res %>% rownames_to_column("probe") %>% filter(P.Value < (0.05/ n()))
write.csv(tmp$probe,row.names=F)
```
### PRSxAge interaction

@TODO FILL THIS IN 
### PRSxSex interaction

```{r}
design_sex <- model.matrix(~1+SCORE1_AVG*sex+age+plate+V3+V4+V5, data=reg_all_data[na_row,])
sex_fit <- lmFit(betas_sub[,na_row],design_sex)
sex_fit <- eBayes(sex_fit)
sex_res <- topTable(sex_fit, number = Inf,coef=9)
sex_main_res <- topTable(sex_fit, number = Inf,coef=c(2,9))
```
```{r}
ggplot(sex_res,aes(logFC,-log10(P.Value)))+ geom_point()
ggplot(sex_res,aes(logFC,-log10(sex_main_res[rownames(sex_res),]$P.Value)))+ geom_point()
```
#### Are these probes different than those with genetic main effect
```{r}
prs_res %>% rownames_to_column("probe") %>% filter(P.Value < (0.05/ n()))
sex_main_res%>% rownames_to_column("probe") %>% filter(P.Value < (0.05/ n()))
sex_res%>% rownames_to_column("probe") %>% filter(P.Value < (0.05/ n()))
ggplot(reg_all_data,aes(SCORE1_AVG,betas_sub["cg20163478",],color = sex)) + geom_point() + geom_smooth(method="lm") + labs(y = "cg20163478 Methylation", x="PD PRS")
```

### PRSxExposure Interaction
Loading in multiple imputation data
```{r}
library(mice)
pest_missing <- read.csv("/home1/NEURO/SHARE_DECIPHER/TERRE_pesticides/pesticides.csv")
#pest_missing <- pest_missing[pest_missing$num %in%colnames(betas_sub), ]
pest_imputed <- read.csv("/home1/NEURO/SHARE_DECIPHER/TERRE_pesticides/pesticides_imputed.csv")
#pest_imputed <- pest_imputed[pest_imputed$num %in%colnames(betas_sub), ]
pre_mids <-rbind(cbind(data.frame(X_imputation_=0),pest_missing)[,colnames(pest_imputed)],pest_imputed)
pre_mids <- pre_mids[pre_mids$num %in% colnames(betas_sub),]
pest_mids <- as.mids(pre_mids, .imp="X_imputation_",.id="num")
##@TODO RE-RUN ANALYSES AND AGGREGATE WITH RUBIN'S METHOD WITH mice PACKAGE
betas_pest <- betas_sub[,rownames(pest_mids$data)]
```
```{r}
exposures <- colnames(pest_mids$data)[colMeans(pest_mids$data,na.rm = T) > 0.025]
pesticide_interactions <- lapply(exposures, function(pesticide) fread(sprintf("%s_PRS_interaction.txt.gz",pesticide)))
names(pesticide_interactions) <- exposures
pesticides_df <- rbindlist(pesticide_interactions,idcol="exposure") %>% group_by(exposure) %>% mutate(FDR = p.adjust(P,method = "BH"))
```
### PRS exposure interaction with Random Forest Imputation
```{r}
library(missForest)
to_impute <- pest_missing[,-c(1)] %>% mutate_all(~factor(.))
pest_forest <- missForest(to_impute)
pest_forest$OOBerror
pest_forest_imputed <- cbind(num=pest_missing$num,pest_forest$ximp %>% mutate_all(~as.numeric(as.character(.))))
reg_w_all_pesticide <- reg_all_data %>% left_join(pest_forest_imputed,by = c("patient"="num"))  %>% filter(!is.na(ins))

reg_pre_impute <- reg_all_data %>% select(patient,SCORE1_AVG,V3,V4,V5,age,men,MMS) %>% left_join(pest_missing %>% mutate_all(~factor(.)), by=c("patient"="num"))
pest_forest_w_vars <- missForest(reg_pre_impute[,-c(1:5)])
reg_all_vars <- cbind(reg_pre_impute[,1:8],pest_forest_w_vars$ximp[,-c(1:3)] %>% mutate_all(~as.numeric(as.character(.)))) %>%filter(patient %in% reg_w_all_pesticide$patient)
pest_forest_w_vars$OOBerror

```

Run regressions:
```{r}
betas_pest <- betas_sub[,as.character(reg_w_all_pesticide$patient)]
all(reg_w_all_pesticide$patient == colnames(betas_pest))
exposures <- colnames(pest_forest$ximp)[colMeans(reg_w_all_pesticide[,.SD,.SDcols=colnames(pest_forest$ximp)]) > 0.025]
# exposures_vars <- colnames(pest_forest_w_vars$ximp[,-c(1:3)])[colMeans(reg_all_vars[,.SD,.SDcols=colnames(pest_forest_w_vars$ximp[,-c(1:3)])]) > 0.04]
res <- parallel::mclapply(
    exposures,
    function(x){
      design <- substitute(model.matrix(~E*SCORE1_AVG + sex + age + V3 + V4 + V5,reg_w_all_pesticide),list(E=as.name(x)))
      fit <- lmFit(betas_pest,eval(design))
      eb <- eBayes(fit)
      topTable(eb,coef=9,number = Inf)
    },
    mc.cores = 16)
names(res) <- as.character(exposures)

# res_vars <- parallel::mclapply(
#     exposures_vars,
#     function(x){
#       design <- substitute(model.matrix(~E*SCORE1_AVG + men + age + V3 + V4 + V5,reg_all_vars),list(E=as.name(x)))
#       fit <- lmFit(betas_pest,eval(design))
#       eb <- eBayes(fit)
#       topTable(eb,coef=9,number = Inf)
#     },
#     mc.cores = 16)
# names(res_vars) <- as.character(exposures_vars)

all_pest_prs_df <- rbindlist(lapply(res,function(df) rownames_to_column(df,var = "probe")),idcol="exposure") %>% filter(adj.P.Val < 0.25) %>% left_join(annot,by=c("probe"="Name"))
(prs_int_hits <- all_pest_prs_df %>% arrange(P.Value) %>% filter(adj.P.Val < 0.05) %>% select(exposure,probe,logFC,P.Value,adj.P.Val, UCSC_RefGene_Name))
# all_pest_prs_vars_df <- rbindlist(lapply(res_vars,function(df) rownames_to_column(df,var = "probe")),idcol="exposure") %>% filter(adj.P.Val < 0.25) %>% left_join(annot,by=c("probe"="Name"))
# all_pest_prs_vars_df %>% arrange(P.Value)
prs_fong_s_cu <- rbindlist(lapply(res,function(df) rownames_to_column(df,var = "probe")),idcol="exposure") %>% filter(exposure %in% c("fong","s","cu")) %>% left_join(annot,by=c("probe"="Name"))
print(prs_int_hits)
```
```{r}
to_plot <- prs_fong_s_cu%>% mutate(exposure=recode(exposure,cu="Copper",fong="Fungicides", s="Sulphur"))
ggplot(to_plot,aes(logFC,-log10(P.Value),color = exposure)) + geom_point() +geom_text_repel(data = subset(to_plot, adj.P.Val < 0.05),mapping=aes(logFC,-log10(P.Value),label=UCSC_RefGene_Name)) + labs(color="Exposure",x=quote(Delta~beta~Methylation))+ theme_minimal()
```

```{r}
methy_tmp <- betas_pest["cg13975855",]
ggplot(reg_w_all_pesticide,aes(SCORE1_AVG,methy_tmp,color = ifelse(fong==1 | cu==1 | s==1,"Exposed","Not Exposed"))) + geom_point() + geom_smooth(method="lm",se = F) + facet_grid(~PD) + labs(x= "PRS for PD", color="Copper, sulphur, or fungicide", y= quote(beta~Methylation~of~cg13975855)) + theme_minimal()
reg_w_all_pesticide <- reg_w_all_pesticide %>% mutate(combined = ifelse(fong==1 | cu==1 | s==1,"Exposed","Not Exposed"))
ggplot(reg_w_all_pesticide,aes(SCORE1_AVG,methy_tmp)) + geom_smooth(method="lm",se = F) + geom_point(data=reg_w_all_pesticide,aes(color=PD)) +facet_grid(~combined) + labs(x= "Polygenic risk score for PD", y= quote(beta~Methylation~of~cg13975855), color="PD status") + theme_minimal()
ggplot(reg_w_all_pesticide,aes(SCORE1_AVG,methy_tmp,color = cu==1)) + geom_point() + geom_smooth(method="lm")
ggplot(reg_w_all_pesticide,aes(SCORE1_AVG,methy_tmp,color = s==1)) + geom_point() + geom_smooth(method="lm")
```

# CASE and CONTROL only analysis
## Pre-check that PRS is independent of exposure

```{r}
prs_vs_pest <- lapply(exposures, function(x) t.test(reg_w_all_pesticide$SCORE1_AVG ~ unlist(reg_w_all_pesticide[,..x]))$p.value)
names(prs_vs_pest) <- exposures
prs_vs_pest %>% stack() %>% arrange(values)
exposures_ind <- names(prs_vs_pest[prs_vs_pest > 0.05])
```
Checking correlation between exposures:
```{r}
library(pheatmap)
cor(reg_w_all_pesticide[,..exposures_ind]) %>% 
  pheatmap(main="Correlation between exposures")
  #melt() %>%
  #ggplot(aes(Var1,Var2,fill=value)) + geom_tile() + scale_fill_viridis_c()+ 
  #theme(axis.text.x=element_text(angle=90,hjust=1,vjust=1))
cor(reg_w_all_pesticide[PD == "CASE",..exposures_ind]) %>% 
  pheatmap(main="Correlation between exposures cases only")
cor(reg_w_all_pesticide[PD == "CONTROL",..exposures_ind]) %>% 
  pheatmap(main="Correlation between exposures controls only")
```

## TERRE CASE ONLY ANALYSIS
```{r}
betas_pest_case <- betas_pest[,reg_w_all_pesticide$PD == "CASE"]
reg_w_all_pesticide_case <- reg_w_all_pesticide %>% filter(PD == "CASE")
all(reg_w_all_pesticide$patient == colnames(betas_pest))
exposures <- colnames(pest_forest$ximp)[colMeans(reg_w_all_pesticide_case[,.SD,.SDcols=colnames(pest_forest$ximp)]) > 0.025 & colnames(pest_forest$ximp) %in% exposures_ind]
res_case <- parallel::mclapply(
    exposures,
    function(x){
      design <- substitute(model.matrix(~E*SCORE1_AVG + sex + age + V3 + V4 + V5,reg_w_all_pesticide_case),list(E=as.name(x)))
      fit <- lmFit(betas_pest_case,eval(design))
      eb <- eBayes(fit)
      topTable(eb,coef=9,number = Inf)
    },
    mc.cores = 16)
names(res_case) <- as.character(exposures)
all_pest_prs_case_df <- rbindlist(lapply(res_case,function(df) rownames_to_column(df,var = "probe")),idcol="exposure") %>% filter(adj.P.Val < 0.05) %>% left_join(annot,by=c("probe"="Name"))
all_pest_prs_case_df %>% arrange(P.Value) %>% select(exposure,probe,logFC,AveExpr,t,P.Value,adj.P.Val,B,CHR,MAPINFO,UCSC_RefGene_Name) %>% View()
```
```{r}
to_plot <- rbindlist(lapply(res_case,function(df) rownames_to_column(df,var = "probe")),idcol="exposure") %>% filter(exposure %in% unique(all_pest_prs_case_df$exposure[order(all_pest_prs_case_df$adj.P.Val)])[1:5]) %>% left_join(annot,by=c("probe"="Name")) %>% mutate(gene=gsub(";.*","",UCSC_RefGene_Name))
ggplot(to_plot,aes(logFC,-log10(P.Value),color = exposure)) + geom_point() +geom_text_repel(data = subset(to_plot, adj.P.Val < 0.05),mapping=aes(logFC,-log10(P.Value),label=gene)) + labs(color="Exposure",x=quote(Delta~beta~Methylation))+ theme_minimal()
```


# TERRE control ONLY ANALYSIS
```{r}
betas_pest_control <- betas_pest[,reg_w_all_pesticide$PD == "CONTROL"]
reg_w_all_pesticide_control <- reg_w_all_pesticide %>% filter(PD == "CONTROL")
all(reg_w_all_pesticide$patient == colnames(betas_pest))
exposures <- colnames(pest_forest$ximp)[colMeans(reg_w_all_pesticide_control[,.SD,.SDcols=colnames(pest_forest$ximp)]) > 0.025& colnames(pest_forest$ximp) %in% exposures_ind]
res_control <- parallel::mclapply(
    exposures,
    function(x){
      design <- substitute(model.matrix(~E*SCORE1_AVG + sex + age + V3 + V4 + V5,reg_w_all_pesticide_control),list(E=as.name(x)))
      fit <- lmFit(betas_pest_control,eval(design))
      eb <- eBayes(fit)
      topTable(eb,coef=9,number = Inf)
    },
    mc.cores = 16)
names(res_control) <- as.character(exposures)
all_pest_prs_control_df <- rbindlist(lapply(res_control,function(df) rownames_to_column(df,var = "probe")),idcol="exposure") %>% filter(adj.P.Val < 0.05) %>% left_join(annot,by=c("probe"="Name"))
all_pest_prs_control_df %>% arrange(P.Value) %>% select(exposure,probe,logFC,AveExpr,t,P.Value,adj.P.Val,B,CHR,MAPINFO,UCSC_RefGene_Name) %>% View()
```
```{r}
to_plot <- rbindlist(lapply(res_control,function(df) rownames_to_column(df,var = "probe")),idcol="exposure") %>% filter(exposure %in% unique(all_pest_prs_control_df$exposure[order(all_pest_prs_control_df$adj.P.Val)])[1:5]) %>% left_join(annot,by=c("probe"="Name"))%>% mutate(gene=gsub(";.*","",UCSC_RefGene_Name))
ggplot(to_plot,aes(logFC,-log10(P.Value),color = exposure)) + geom_point() +geom_text_repel(data = subset(to_plot, adj.P.Val < 0.05),mapping=aes(logFC,-log10(P.Value),label=gene)) + labs(color="Exposure",x=quote(Delta~beta~Methylation))+ theme_minimal()
```
# DIGPD PRS Analysis
Check SNPs in set of variants:

```{r}
bim <- fread("~/genotype_qc/DIGPD_merged_post_imputation_QC/raw_data.imputed.r2_90.hard_call.maf_mismatch.bim", header=F)
sum(risk_scores$SNP %in% bim$V2)
```

Generate PRS using `plink2 --score`:
```{bash}
~/plink2 --bfile ~/genotype_qc/DIGPD_merged_post_imputation_QC/raw_data.imputed.r2_90.hard_call.maf_mismatch\
  --score pd_prs_file.txt list-variants\
  --out digpd_prs

~/plink2 --bfile ~/genotype_qc/DIGPD_case_QC/all_imputed_r2_30_rsid\
  --score pd_prs_file.txt list-variants\
  --out digpd_control_prs
  
~/plink2 --bfile ~/genotype_qc/DIGPD_control_QC/all_imputed_r2_30_rsid\
  --score pd_prs_file.txt list-variants\
  --out digpd_case_prs
  
  
```
## Loading in and preparing data
```{r}
digpd_merged_fam <- rbind(fread("~/genotype_qc/DIGPD_case_pre_pca.matched.fam"),fread("~/genotype_qc/DIGPD_control_pre_pca.matched.fam"))
digpd_prs <- fread("digpd_prs.sscore")
digpd_prs$sex <- ifelse(digpd_merged_fam$V5 == 1,"MALE","FEMALE")
digpd_prs$PD <- ifelse(digpd_merged_fam$V6 == 2,"CASE","CONTROL")
digpd_control_prs <- fread("digpd_control_prs.sscore")
digpd_case_prs <- fread("digpd_case_prs.sscore")
strat_prs <- rbind(digpd_control_prs, digpd_case_prs)
digpd_pcs <- fread("~/genotype_qc/DIGPD_merged_post_imputation_QC/raw_data.preimpute.pcs")
digpd_metadata <- fread("/home1/NEURO/SHARE_DECIPHER/digpd_meta_master.csv") %>% select(-PD) %>% mutate(IID = gsub("PAL_[0-9]*_","",IID))
all_digpd_data <- digpd_prs %>% left_join(digpd_metadata, by=c("IID")) %>% left_join(digpd_pcs,by=c("IID"))
```

### PRS score testing

```{r}
ggplot(all_digpd_data,aes(PD,SCORE1_AVG)) + geom_boxplot()+ geom_jitter(height=0, width=0.1)+ labs(x= "Parkinson's Diagnosis", y = "Polygenic Risk Score") +ggtitle(bquote('Nagelkerke Pseudo-'*R^2*':0.0578120' ))
t.test(all_digpd_data$SCORE1_AVG[all_digpd_data$PD == "CASE"],all_digpd_data$SCORE1_AVG[all_digpd_data$PD == "CONTROL"], alternative = "greater") 

```

### PRS DIGPD PD associations
```{r}
all_digpd_data$PD_fixed <- ifelse(all_digpd_data$PD == "CASE",1,0)
fit1 <- glm(PD_fixed ~ SCORE1_AVG + sex  +PC1 + PC2 + PC3, data=all_digpd_data,family="binomial",control = list(maxit = 50))
summary(fit1)
fit1_null <- glm(PD_fixed ~ sex +PC1 + PC2 + PC3, data=all_digpd_data,family="binomial",control = list(maxit = 50))
summary(fit1_null)
nagelkerke(fit1,null=fit1_null)
lin1<- lm(PD_fixed ~ SCORE1_AVG + sex  +PC1 + PC2 + PC3, data=all_digpd_data)
lin2<- lm(PD_fixed ~sex  +PC1 + PC2 + PC3, data=all_digpd_data)
R2 <- (deviance(lin2) - deviance(lin1)) / deviance(lin2)
h2 <- pd_h2(R2,sum(all_digpd_data$PD_fixed)/nrow(all_digpd_data),0.01)
fit2 <- glm(PD_fixed ~ SCORE1_AVG*sex + PC1+PC2+PC3, data = all_digpd_data, family="binomial")
summary(fit2)
fit3 <- glm(PD_fixed~SCORE1_AVG*smoking +sex+ PC1+PC2+PC3, data=all_digpd_data, family="binomial")
summary(fit3)
fit4 <- glm(PD_fixed~SCORE1_AVG*pesticides +sex+ PC1+PC2+PC3, data=all_digpd_data, family="binomial")
summary(fit4)


```
## DIGPD PRS methylation associations
```{r}
load("/home1/NEURO/SHARE_DECIPHER/processed_DNAm_data/DIGPD/DIGPDvisit1_betas_combat.RData")
all_digpd_data$patient <- paste0(all_digpd_data$patient,".visit",all_digpd_data$visite)
digpd_reg_data <- all_digpd_data %>% filter(patient %in% colnames(betas_sub)) %>% mutate(plate=factor(plate))
betas_digpd <- betas_sub[,digpd_reg_data$patient]
```
### including covariates

```{r}
pcs <- prcomp(t(betas_digpd))
```
```{r}
library(pheatmap)
tmp_design <- model.matrix(~0+.,data=digpd_reg_data[,c("race1","plate","birth_cohort")] %>% mutate(race1=factor(race1),birth_cohort = birth_cohort - min(birth_cohort)))
pheatmap(
  cor(pcs$x[,1:20],
      cbind(tmp_design,digpd_reg_data[,c("men","PD","levodopa","Age","pesticides","durevol","smoking","bmi")]) %>% mutate(PD = as.numeric(PD=="CASE")),
    use="pairwise.complete.obs"
  ),
  cluster_rows = F
)
```

### digpd PRS association
```{r}
design_prs <- model.matrix(~1+SCORE1_AVG+Age+men+PC1+PC2+PC3 + birth_cohort, data=digpd_reg_data %>% mutate(race1=factor(race1),birth_cohort = birth_cohort - min(birth_cohort),PD = as.numeric(PD=="CASE")))
prs_fit <- lmFit(betas_digpd,design_prs)
prs_fit <- eBayes(prs_fit)
design_pd <- model.matrix(~1+PD+Age+men+plate+birth_cohort, data=digpd_reg_data %>% mutate(race1=factor(race1),birth_cohort = birth_cohort - min(birth_cohort),PD = as.numeric(PD=="CASE")))
pd_fit <- lmFit(betas_digpd,design_pd)
pd_fit <- eBayes(pd_fit)
pd_res <- topTable(pd_fit,coef=2,number = Inf)
prs_res <- topTable(prs_fit,coef=2,number = Inf)
```


```{r}
library(ggrepel)
annot <- fread("~/MethylationEPIC_v-1-0_B4.csv",skip = 7)
pd_annot_digpd <- pd_res %>%rownames_to_column("probe") %>% left_join(annot,by=c("probe"="Name"))%>% mutate(gene = gsub(";.*","",UCSC_RefGene_Name)) %>% mutate(gene= ifelse(gene == "",probe,gene))
prs_annot_digpd <- prs_res %>%rownames_to_column("probe") %>% left_join(annot,by=c("probe"="Name"))%>% mutate(gene = gsub(";.*","",UCSC_RefGene_Name)) %>% mutate(gene= ifelse(gene == "",probe,gene))

ggplot(pd_annot_digpd,aes(logFC,-log10(P.Value)))+ geom_point() + geom_text_repel(data=pd_annot_digpd %>% filter(abs(logFC) > 0.03 & adj.P.Val < 0.05) %>% top_n(25,-adj.P.Val),mapping=aes(logFC,-log10(P.Value),label=gene))  + geom_point(data=pd_annot_digpd %>% filter(abs(logFC) > 0.03 & adj.P.Val < 0.05), color = "red",mapping=aes(logFC,-log10(P.Value))) +labs(x = quote(Delta~beta~Methylation)) + theme_minimal() + geom_hline(linetype="dashed",yintercept = min(-log10(pd_annot_digpd$P.Value[pd_annot_digpd$adj.P.Val <0.05]))) + geom_vline(linetype="dashed",xintercept = 0.03) + geom_vline(linetype="dashed",xintercept = -0.03)

ggplot(prs_annot_digpd,aes(logFC,-log10(P.Value)))+ geom_point() + geom_text_repel(data=subset(prs_annot_digpd,adj.P.Val < 0.05& abs(logFC) > 0.03),mapping=aes(logFC,-log10(P.Value),label=gene))  + geom_point(data=subset(prs_annot_digpd,adj.P.Val < 0.05 & abs(logFC) > 0.03), color = "red",mapping=aes(logFC,-log10(P.Value))) +labs(x = quote(Delta~beta~Methylation)) + theme_minimal() + geom_hline(linetype="dashed",yintercept = min(-log10(prs_annot_digpd$P.Value[prs_annot_digpd$adj.P.Val <0.05]))) + geom_vline(linetype="dashed",xintercept = 0.03) + geom_vline(linetype="dashed",xintercept = -0.03)
```
### checking reason for difference in hits between DIGPD and Terre
Suspicious inflation here:
```{r}
qq(pd_annot_digpd$P.Value,main="DIGPD Parkinson's EWAS")
qq(pd_annot$P.Value,main="Terre Parkinson's EWAS")
qq(prs_annot_digpd$P.Value,main="DIGPD Parkinson's PRS EWAS")
qq(prs_annot$P.Value,main="Terre Parkinson's PRS EWAS")
```

```{r}
library(qqman)
manhattan(prs_annot_digpd %>% mutate(P=P.Value,SNP=gene, CHR=as.numeric(recode(CHR,X="23",Y="24",MT="26")),BP=MAPINFO),chrlabs = c(1:22,"X","Y"),annotateTop = F,annotatePval = 5e-8)
```
```{r}
qq(pd_annot_digpd$P.Value,main="DIGPD Parkinson's EWAS")
```

