---
title: "Analysis of PRS interactions"
output: html_notebook
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
library(ggplotify)
library(grid)
library(gridExtra)
library(data.table)
library(qqman)
library(limma)
knitr::opts_chunk$set(echo = TRUE)
```
# Pesticides
## Checking inflation for prsxpest summary stats
Before I suspect I was combining effects from limma incorrectly. So I re-rolled the regressions using mice and lm as the mice package seems to be intended to be used. Let's check inflation now within each pesticide:
### Cross-sex
```{r}
pesticide_interactions <- fread("prs_interaction_result_prsice.txt.gz")
plots <- list()
for (cur_env in unique(pesticide_interactions$env)) {
  plots[[cur_env]] <- as.grob(~ qq(pesticide_interactions[pesticide_interactions$env == cur_env, ]$GxEp, main = cur_env))
}
```
```{r, fig.width=11,fig.height=12}
do.call(plot_grid, plots)
```

#### Filtering sumstats based on within-exposure FDR
```{r}
manifest <- fread("~/MethylationEPIC_v-1-0_B4.csv", skip = 7, fill = TRUE)
pesticide_interactions[, c("GxEq", "Eq") := .(p.adjust(GxEp, method = "fdr"), p.adjust(Ep, method = "fdr")), by = "env"]
(sig_interactions <- pesticide_interactions[GxEq < 0.25][order(GxEq)])
(sig_E <- pesticide_interactions[Eq < 0.05][order(Eq)])
```
```{r}
merge(sig_interactions, manifest, by.x = "cpg", by.y = "Name")[, .(cpg, env, GxEp, GxEq, Relation_to_UCSC_CpG_Island, UCSC_RefGene_Name)][order(GxEq)]

merge(sig_interactions, manifest, by.x = "cpg", by.y = "Name")[, .(cpg, env, GxEp, GxEq, Relation_to_UCSC_CpG_Island, UCSC_RefGene_Name)][order(GxEq)] %>% write.csv(quote = F)
```

### Males
```{r}
pesticide_interactions <- fread("prs_interaction_result_prsice_male.txt.gz")
plots <- list()
for (cur_env in unique(pesticide_interactions$env)) {
  plots[[cur_env]] <- as.grob(~ qq(pesticide_interactions[pesticide_interactions$env == cur_env, ]$GxEp, main = cur_env))
}
```
```{r, fig.width=11,fig.height=12}
do.call(plot_grid, plots)
```

#### Filtering sumstats based on within-exposure FDR @TODO won't run
```{r}
pesticide_interactions[, c("GxEq", "Eq") := .(p.adjust(GxEp, method = "fdr"), p.adjust(Ep, method = "fdr")), by = "env"]
sig_interactions <- pesticide_interactions[GxEq < 0.25][order(GxEq)]
sig_E <- pesticide_interactions[Eq < 0.05][order(Eq)]
```
### Females
```{r}
pesticide_interactions <- fread("prs_interaction_result_prsice_female.txt.gz")
plots <- list()
for (cur_env in unique(pesticide_interactions$env)) {
  plots[[cur_env]] <- as.grob(~ qq(pesticide_interactions[pesticide_interactions$env == cur_env, ]$GxEp, main = cur_env))
}
```
```{r, fig.width=11,fig.height=12}
do.call(plot_grid, plots)
```

#### Filtering sumstats based on within-exposure FDR
```{r}
pesticide_interactions[, c("GxEq", "Eq") := .(p.adjust(GxEp, method = "fdr"), p.adjust(Ep, method = "fdr")), by = "env"]
(sig_interactions <- pesticide_interactions[GxEq < 0.25][order(GxEq)])
(sig_E <- pesticide_interactions[Eq < 0.05][order(Eq)])
```

## Pesticide interactions within GRS hits

```{r}
cross <- fread("prs_interaction_results_cross.txt")[,c("GxEq", "Eq") := .(p.adjust(GxEp, method = "fdr"), p.adjust(Ep, method = "fdr"))]
male <- fread("prs_interaction_results_male.txt")[,c("GxEq", "Eq") := .(p.adjust(GxEp, method = "fdr"), p.adjust(Ep, method = "fdr"))]
female <- fread("prs_interaction_results_female.txt")[,c("GxEq", "Eq") := .(p.adjust(GxEp, method = "fdr"), p.adjust(Ep, method = "fdr"))]
cross %>% arrange(GxEq) %>% filter(GxEq < 0.25)
cross %>% arrange(GxEq) %>% filter(Eq < 0.25)

female %>% arrange(GxEq) %>% filter(GxEq < 0.25)
female %>% arrange(GxEq) %>% filter(Eq < 0.25)

female %>% arrange(GxEq) %>% filter(GxEq < 0.25)
female %>% arrange(GxEq) %>% filter(Eq < 0.25)
```

## Interaction effects on non-pesticide variables
### Cross-sex
```{r}
#loading
load("/home1/NEURO/SHARE_DECIPHER/processed_DNAm_data/2022/TERRE_processed_2022/1-TERRE_RG_filtered.RData") # PD_RG_filtered
covars <- fread("/home1/NEURO/SHARE_DECIPHER/terre_meta_master.csv")
covars$IID <- gsub("_PAE.*", "", covars$IID)
mvals_combat <- minfi::getM(PD_RG_filtered)
prs_mat <- fread("prsice_data/TERRE_PRSice.all_score")

#filtering
covars <- covars[patient %in% colnames(mvals_combat) & IID %in% prs_mat$IID]
prs_mat <- prs_mat[IID %in% covars$IID]
mvals_combat <- mvals_combat[,colnames(mvals_combat) %in% covars$patient]

#matching
covars <- covars[match(patient,colnames(mvals_combat))]
colnames(mvals_combat) <- covars$IID[match(covars$patient,colnames(mvals_combat))]
prs_mat <- prs_mat[match(IID,colnames(mvals_combat))]


computed_covars <- fread("prsice_data/TERRE.covariate")[IID %in% colnames(mvals_combat)]


all(covars$IID == computed_covars$IID)
covars <- cbind(covars[,-c("age","men")],prs_mat[,.(SCORE1_AVG=`Pt_5e-08`)],computed_covars)
covars <- covars[,.(SCORE1_AVG,age,alcohol=ifelse(alcohol1 > 2,1,0), smoking=ifelse(smoking > 1 ,1,0),GPC1=V3,GPC2=V4,GPC3=V5,PC1,PC2,PC3,PC4,PC5,PC6,PC7,PC8,PC9,PC10,head_trauma=head_trauma_loc,sex)]

all(covars$IID == prs_mat$IID)
all(prs_mat$IID == colnames(mvals_combat))

```

```{r}
corrplot::corrplot(cor(model.matrix(~.-1,data=covars)),type = "upper",diag = FALSE)
```

## Interaction function
```{r}
base_model <- ~ sex + age + GPC1 + GPC2 + GPC3 + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10
fit_interaction  <- function(methy, term){
  ix <- which(!is.na(covars[,..term][[1]]))
  interaction_term <- paste0("SCORE1_AVG*",term)
  design <- model.matrix(update(base_model,paste("~ . +", interaction_term)), data=covars[ix])
  fit <- lmFit(methy[,ix], design)
  fit <- eBayes(fit)
  if(paste0(term,":SCORE1_AVG") %in% colnames(fit$coefficients)){
    res <- topTable(fit, number = Inf, coef = paste0(term,":SCORE1_AVG"))
  }else{
    res <- topTable(fit, number = Inf, coef = paste0("SCORE1_AVG:",term))
  }
  main_res <- topTable(fit, number = Inf, coef = term)
  return(list(interaction_effect=res,main=main_res))
}

sex_res <- fit_interaction(mvals_combat,"sex")
age_res <- fit_interaction(mvals_combat,"age")
head_trauma_res <- fit_interaction(mvals_combat,"head_trauma")
alcohol_res <- fit_interaction(mvals_combat,"alcohol")
smoking_res <- fit_interaction(mvals_combat,"smoking")
save(list=c("sex_res","age_res","head_trauma_res","alcohol_res","smoking_res"),file="non-pesticide-interaction-results.RData")

```
