---
title: "PRS Analysis"
output: html_notebook
---

```{r setup, include=FALSE}
library(DMRcate)
library(limma)
library(minfi)
library(rcompanion)
library(readxl)
library(tidyverse)
library(ggpubr)
library(glue)
library(data.table)
library(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
knitr::opts_chunk$set(echo = TRUE)
```
# PD risk score data and matching
```{r, eval=FALSE}
risk_scores <- read_xlsx("~/pd_risk_score/Supplementary materials/Table S2. Detailed summary statistics on all nominated risk variants, known and novel_.xlsx") %>%
  mutate(CHR = as.numeric(CHR), COUNTED = toupper(`Effect allele`), ALT = toupper(`Other allele`), POS = BP) %>%
  filter(!is.na(CHR))
risk_scores %>%
  select(SNP, COUNTED, `Beta, all studies`) %>%
  write_delim("pd_prs_file.txt", col_names = FALSE)
```

# TERRE PRS Analysis
Generate PRS using `plink2 --score`:
```{bash, eval=FALSE}
~/plink2 --bfile ~/genotype_qc/TERRE_QC/raw_data.imputed.r2_30.hard_call.maf_mismatch\
  --score pd_prs_file.txt\
  --out terre_prs
```
## Loading in and preparing data
```{r}
terre_prs <- fread("terre_prs.sscore")
terre_pcs <- fread("~/genotype_qc/TERRE_QC/raw_data.geno.maf.mind.sex_check.het_filter.ibd_filter.eigenvec")
terre_metadata <- read.csv("/home1/NEURO/SHARE_DECIPHER/terre_meta_master.csv") %>%
  mutate(
    FID = gsub("(PAE_[0-9]*_[1-9]*)_.*", "\\1", FID),
    IID = gsub("(PAE_[0-9]*_[1-9]*)_.*", "\\1", IID)
  )
all_data <- terre_prs %>%
  left_join(terre_metadata, by = c("IID")) %>%
  left_join(terre_pcs, by = c("IID" = "V1"))
```
## Association analysis
```{r}
pd_h2 <- function(R2O, case_prop, prevalence = 0.22) {
  K <- prevalence
  P <- case_prop
  thd <- -1 * qnorm(K, 0, 1)
  zv <- dnorm(thd) # z (normal density)
  mv <- zv / K # mean liability for case
  theta <- mv * (P - K) / (1 - K) * (mv * (P - K) / (1 - K) - thd) # θ in equation
  cv <- K * (1 - K) / zv^2 * K * (1 - K) / (P * (1 - P)) # C in
  return(R2O * cv / (1 + R2O * theta * cv))
}
# R2
fit1 <- glm(PD ~ SCORE1_AVG + men + age + V3 + V4 + V5, data = all_data, family = "binomial")
summary(fit1)
fit1_null <- glm(PD ~ men + age + V3 + V4 + V5, data = all_data, family = "binomial")
summary(fit1_null)
(cross_r2 <- nagelkerke(fit1, null = fit1_null))
# liability R2
lin1 <- lm(PD ~ SCORE1_AVG + men + age + V3 + V4 + V5, data = all_data)
lin2 <- lm(PD ~ men + age + V3 + V4 + V5, data = all_data)
R2 <- (deviance(lin2) - deviance(lin1)) / deviance(lin2)
h2 <- pd_h2(R2, sum(all_data$status) / nrow(all_data), 5 / 1000)
R2
h2

# Sex stratified R2
fit1_male <- glm(PD ~ SCORE1_AVG + age + V3 + V4 + V5, data = all_data[men == 1, ], family = "binomial")
summary(fit1_male)
fit1_male_null <- glm(PD ~ age + V3 + V4 + V5, data = all_data[men == 1, ], family = "binomial")
summary(fit1_male_null)
(male_r2 <- nagelkerke(fit1_male, null = fit1_male_null))
fit1_female <- glm(PD ~ SCORE1_AVG + age + V3 + V4 + V5, data = all_data[men == 0, ], family = "binomial")
summary(fit1_female)
fit1_female_null <- glm(PD ~ age + V3 + V4 + V5, data = all_data[men == 0, ], family = "binomial")
summary(fit1_female_null)
(female_r2 <- nagelkerke(fit1_female, null = fit1_female_null))
# Additional tests
fit2 <- glm(PD ~ SCORE1_AVG * men + V3 + V4 + V5, data = all_data, family = "binomial")
summary(fit2)
fit3 <- glm(PD ~ SCORE1_AVG * smoking + V3 + V4 + V5, data = all_data, family = "binomial")
summary(fit3)
fit4 <- glm(PD ~ SCORE1_AVG * pesticides + V3 + V4 + V5, data = all_data, family = "binomial")
summary(fit4)

fit_null <- glm(PD ~ SCORE1_AVG, data = all_data, family = "binomial")
summary(fit_null)
```

```{r, fig.width=10}
r2_labels <- c(
  round(male_r2$Pseudo.R.squared.for.model.vs.null[3], 4),
  round(female_r2$Pseudo.R.squared.for.model.vs.null[3], 4),
  0.0611
)
h2_labels <- c(
  round(pd_h2(male_r2$Pseudo.R.squared.for.model.vs.null[3], sum(all_data$status) / nrow(all_data), 5 / 1000), 4),
  round(pd_h2(female_r2$Pseudo.R.squared.for.model.vs.null[3], sum(all_data$status) / nrow(all_data), 5 / 1000), 4),
  0.0241
)

to_plot <- all_data %>%
  mutate(
    PD = ifelse(PD, "CASE", "CONTROL"),
    Sex = ifelse(men == 1, "Male", "Female")
  ) %>%
  mutate(Sex = factor(Sex, levels = c("Cross-sex", "Male", "Female")))
cross_data <- all_data %>% mutate(PD = ifelse(PD, "CASE", "CONTROL"))
cross_data$Sex <- "Cross-sex"
to_plot <- rbind(cross_data, to_plot)

ggboxplot(
  to_plot,
  x = "PD",
  y = "SCORE1_AVG",
  color = "Sex",
  add = "jitter"
) +
  facet_grid(~Sex, labeller = label_bquote(cols = .(as.character(Sex)) ~ R^2 ~ ": " ~ .(r2_labels[Sex]) ~ ";" ~ h^2 ~ ": " ~ .(h2_labels[Sex]))) +
  labs(y = "Polygenic Risk Score", x = "Parkinson's Diagnosis") +
  scale_color_manual(values = c("gray", "light blue", "pink"))

ggboxplot(
  to_plot,
  x = "Sex",
  y = "SCORE1_AVG",
  color = "Sex",
  add = "jitter"
) +
  stat_compare_means() +

  labs(y = "Polygenic Risk Score", x = "Parkinson's Diagnosis") +
  scale_color_manual(values = c("Male" = "light blue", "Female" = "pink"))
```

## PRS vs DNAm

```{r}
# load("/home1/NEURO/SHARE_DECIPHER/processed_DNAm_data/TERRE/TERRE_betas_CTC.RData")# terre_betas_CTC
load("/home1/NEURO/SHARE_DECIPHER/processed_DNAm_data/TERRE_processed_2021/betas_combat.RData") # betas_sub
betas_sub <- betas_combat
# load("/home1/NEURO/SHARE_DECIPHER/processed_DNAm_data/TERRE/TERRE_funnorm_processed/betas_combat.RData")
# betas_sub<- betas_combat
```


Let's check how the data looks for a random subject:
```{r}
ggplot(betas_sub[, c(1, 2, 3, 4, 5)] %>% as.data.table(keep.rownames = T) %>% melt(id.vars = "rn", value.name = "betas", variable.name = "subject"), aes(betas, color = subject)) +
  geom_density()
```

### Match DNA, PRS, and metadata
```{r}
reg_all_data <- all_data[na.omit(match(colnames(betas_sub), all_data$patient)), ]
reg_all_data$sex <- ifelse(reg_all_data$men, "Male", "Female")
reg_all_data$PD <- ifelse(reg_all_data$PD, "CASE", "CONTROL")
betas_sub <- betas_sub[, colnames(betas_sub) %in% reg_all_data$patient]
all(colnames(betas_sub) == reg_all_data$patient)
```
### methylation pcs
```{r}
library(RSpectra)
compute_pcs <- function(A, k = 10) {
  n <- nrow(A)
  p <- ncol(A)
  A <- scale(A, center = TRUE, scale = FALSE)
  s <- svds(A, k, nu = 0, nv = k)
  j <- seq_len(k)
  s$d <- s$d / sqrt(max(1, n - 1))

  dimnames(s$v) <- list(colnames(A), paste0("PC", j))
  r <- list(sdev = s$d, rotation = s$v)
  class(r) <- "prcomp"
  r
}
res <- compute_pcs(betas_sub)
PCs <- res$rotation
```
### TERRE PRS regression vs methylation
```{r}
if (!"PC1" %in% colnames(reg_all_data)) reg_all_data <- cbind(reg_all_data, as.data.frame(PCs))
```

#### Differential methylation with PRS vs PD
```{r}
na_row <- !apply(reg_all_data[, c("SCORE1_AVG", "age", "plate", "sex")], 1, function(x) any(is.na(x)))
design_prs <- model.matrix(~ 1 + SCORE1_AVG + age + plate + sex + V3 + V4 + V5 + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = reg_all_data[na_row, ])
prs_fit <- lmFit(betas_sub[, na_row], design_prs)
prs_fit <- eBayes(prs_fit)
# Cases only
case_select <- na_row & reg_all_data$PD == "CASE"
design_prs_case <- model.matrix(~ 1 + SCORE1_AVG + age + plate + sex + V3 + V4 + V5 + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = reg_all_data[case_select, ])
prs_fit_case <- lmFit(betas_sub[, case_select], design_prs_case)
prs_fit_case <- eBayes(prs_fit_case)
# Controls only
control_select <- na_row & reg_all_data$PD == "CONTROL"
design_prs_control <- model.matrix(~ 1 + SCORE1_AVG + age + plate + sex + V3 + V4 + V5 + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = reg_all_data[control_select, ])
prs_fit_control <- lmFit(betas_sub[, control_select], design_prs_control)
prs_fit_control <- eBayes(prs_fit_control)

# Males
male_select <- na_row & reg_all_data$sex == "Male"
design_prs_male <- model.matrix(~ 1 + SCORE1_AVG + age + plate + V3 + V4 + V5 + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = reg_all_data[male_select, ])
prs_fit_male <- lmFit(betas_sub[, male_select], design_prs_male)
prs_fit_male <- eBayes(prs_fit_male)

# Females
female_select <- na_row & reg_all_data$sex == "Female"
design_prs_female <- model.matrix(~ 1 + SCORE1_AVG + age + plate + V3 + V4 + V5 + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = reg_all_data[female_select, ])
prs_fit_female <- lmFit(betas_sub[, female_select], design_prs_female)
prs_fit_female <- eBayes(prs_fit_female)

design_pd <- model.matrix(~ 1 + PD + age + plate + sex + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = reg_all_data[na_row, ])
pd_fit <- lmFit(betas_sub[, na_row], design_pd)
pd_fit <- eBayes(pd_fit)
pd_res <- topTable(pd_fit, coef = 2, number = Inf)
prs_res <- topTable(prs_fit, coef = 2, number = Inf)
control_res <- topTable(prs_fit_control, coef = 2, number = Inf)
case_res <- topTable(prs_fit_case, coef = 2, number = Inf)
male_res <- topTable(prs_fit_male, coef = 2, number = Inf)
female_res <- topTable(prs_fit_female, coef = 2, number = Inf)
```
##### DMR analysis

```{r}
prs_annotated <- cpg.annotate(datatype = "array", object = betas_sub[, na_row], analysis.type = "differential", design = design_prs, coef = 2, what = "Beta", arraytype = "EPIC", fdr = 0.05)
prs_dmr_res <- dmrcate(prs_annotated, lambda = 1000, C = 2)

male_prs_annotated <- cpg.annotate(datatype = "array", object = betas_sub[male_select, na_row], analysis.type = "differential", design = design_prs, coef = 2, what = "Beta", arraytype = "EPIC", fdr = 0.05)
male_prs_dmr_res <- dmrcate(male_prs_annotated, lambda = 1000, C = 2)

female_prs_annotated <- cpg.annotate(datatype = "array", object = betas_sub[female_select, na_row], analysis.type = "differential", design = design_prs, coef = 2, what = "Beta", arraytype = "EPIC", fdr = 0.05)
female_prs_dmr_res <- dmrcate(female_prs_annotated, lambda = 1000, C = 2)

pd_annotated <- cpg.annotate(datatype = "array", object = betas_sub[, na_row], analysis.type = "differential", design = design_pd, coef = 2, what = "Beta", arraytype = "EPIC", fdr = 0.05)
pd_dmr_res <- dmrcate(pd_annotated, lambda = 1000, C = 2)
```
```{r}
S4_to_dataframe <- function(s4obj) {
  nms <- slotNames(s4obj)

  lst <- lapply(nms, function(nm) slot(s4obj, nm))
  as.data.frame(setNames(lst, nms))
}
prs_dmr <- S4_to_dataframe(prs_dmr_res)
male_prs_dmr <- S4_to_dataframe(male_prs_dmr_res)
female_prs_dmr <- S4_to_dataframe(female_prs_dmr_res)
annotation <- minfi::getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
```

##### Plotting
```{r, fig.height=7,fig.width=15}
library(cowplot)
get_dmr_effects <- function(dmr, limma_res) {
  dmr_coord <- str_match_all(as.character(dmr), "(chr.*):([0-9]*)-([0-9]*)")[[1]]
  cpgs <- annotation[annotation$chr == dmr_coord[2] & annotation$pos >= as.numeric(dmr_coord[3]) & annotation$pos <= as.numeric(dmr_coord[4]), ]
  res <- cbind(cpgs, limma_res[cpgs$Name, ])
  dmr_1 <- as.data.frame(res[!is.na(res$logFC), ])
  dmr_methy <- melt(betas_sub[dmr_1$Name, na_row], stringsAsFactors = FALSE)
  to_plot <- merge(dmr_1, dmr_methy, by.x = "Name", by.y = "Var1")
  to_plot$DMR <- dmr
  merge(to_plot, all_data, by.x = "Var2", by.y = "patient")
}
get_dmr_res <- function(dmr, limma_res) {
  dmr_coord <- str_match_all(as.character(dmr), "(chr.*):([0-9]*)-([0-9]*)")[[1]]
  cpgs <- annotation[annotation$chr == dmr_coord[2] & annotation$pos >= as.numeric(dmr_coord[3]) & annotation$pos <= as.numeric(dmr_coord[4]), ]
  res <- cbind(cpgs, limma_res[cpgs$Name, ])
  res$DMR <- dmr
  return(res)
}
plot_dmrs <- function(dmr_res, limma_res, case_control) {
  to_plot <- rbindlist(lapply(dmr_res$coord, function(dmr) get_dmr_effects(dmr, limma_res)))

  p1 <- ggplot(to_plot %>% mutate(SCORE1_AVG = scale(SCORE1_AVG)), aes(pos, value, color = SCORE1_AVG)) +
    geom_point() +
    scale_color_gradient(low = rev(case_control)[1], high = rev(case_control)[2]) +
    facet_wrap(~DMR, scales = "free", nrow = 1) +
    theme_minimal() +
    labs(color = "Normalized PD PRS")
  p2 <- ggplot(to_plot %>% mutate(PD = ifelse(PD == 1, "CASE", "CONTROL")), aes(factor(pos), value, color = PD)) +
    geom_boxplot(position = position_dodge(0.75)) +
    scale_color_manual(values = case_control) +
    facet_wrap(~DMR, scales = "free", nrow = 1) +
    theme_minimal() +
    stat_summary(aes(group = PD), fun = mean, geom = "line") +
    labs(color = "PD status") +
    theme(axis.text.x = element_text(angle = 90))
  print(p1)
  print(p2)
}
plot_dmrs(prs_dmr, prs_res, rev(c("gray80", "gray40")))
plot_dmrs(male_prs_dmr, male_res, rev(c("light blue", "lightblue4")))
plot_dmrs(female_prs_dmr, female_res, rev(c("lightpink", "lightpink4")))
```


```{r, eval=FALSE}
library(ggrepel)
if (!exists("annot")) annot <- fread("~/MethylationEPIC_v-1-0_B4.csv", skip = 7)
pd_annot <- pd_res %>%
  rownames_to_column("probe") %>%
  left_join(annot, by = c("probe" = "Name")) %>%
  mutate(gene = gsub(";.*", "", UCSC_RefGene_Name)) %>%
  mutate(gene = ifelse(gene == "", probe, gene))

prs_annot <- prs_res %>%
  rownames_to_column("probe") %>%
  left_join(annot, by = c("probe" = "Name")) %>%
  mutate(gene = gsub(";.*", "", UCSC_RefGene_Name)) %>%
  mutate(gene = ifelse(gene == "", probe, gene))

ggplot(prs_annot, aes(logFC, -log10(P.Value))) +
  geom_point() +
  geom_text_repel(
    data = prs_annot %>% filter(abs(logFC) > 0.03 & adj.P.Val < 0.05),
    color = "dodgerblue",
    mapping = aes(logFC, -log10(P.Value), label = gene)
  ) +
  geom_point(
    data = subset(prs_annot, adj.P.Val < 0.05 & abs(logFC) > 0.03),
    color = "green",
    mapping = aes(logFC, -log10(P.Value))
  ) +
  labs(x = quote(Delta ~ beta ~ Methylation)) +
  theme_minimal() +
  geom_hline(
    linetype = "dashed",
    yintercept = min(-log10(prs_annot$P.Value[prs_annot$adj.P.Val < 0.05]))
  ) +
  geom_vline(linetype = "dashed", xintercept = 0.03) +
  geom_vline(linetype = "dashed", xintercept = -0.03) +
  ggtitle("Cases and Controls")
# prs_annot[prs_annot$adj.P.Val <0.05,c("gene","CHR","MAPINFO","logFC","P.Value","UCSC_RefGene_Name")] %>% mutate(UCSC_RefGene_Name = sapply(UCSC_RefGene_Name,function(x)split(x)[1]))
prs_highlight <- subset(prs_annot, adj.P.Val < 0.05 & abs(logFC) > 0.03)
case_annot <- case_res %>%
  rownames_to_column("probe") %>%
  left_join(annot, by = c("probe" = "Name")) %>%
  mutate(gene = gsub(";.*", "", UCSC_RefGene_Name)) %>%
  mutate(gene = ifelse(gene == "", probe, gene))
ggplot(case_annot, aes(logFC, -log10(P.Value))) +
  geom_point() +
  geom_text_repel(
    data = subset(case_annot, adj.P.Val < 0.05 & abs(logFC) > 0.03 & !probe %in% prs_highlight$probe),
    color = "dodgerblue",
    mapping = aes(logFC, -log10(P.Value), label = gene)
  ) +
  geom_point(
    data = subset(case_annot, adj.P.Val < 0.05 & abs(logFC) > 0.03 & !probe %in% prs_highlight$probe),
    color = "red",
    mapping = aes(logFC, -log10(P.Value))
  ) +
  geom_text_repel(
    data = case_annot %>% filter(probe %in% prs_highlight$probe), # subset(case_annot,adj.P.Val < 0.05& abs(logFC) > 0.03),
    color = "dodgerblue",
    mapping = aes(logFC, -log10(P.Value), label = gene)
  ) +
  geom_point(
    data = case_annot %>% filter(probe %in% prs_highlight$probe), # subset(case_annot,adj.P.Val < 0.05 & abs(logFC) > 0.03),
    color = "green",
    mapping = aes(logFC, -log10(P.Value))
  ) +
  labs(x = quote(Delta ~ beta ~ Methylation)) +
  theme_minimal() +
  geom_hline(
    linetype = "dashed",
    yintercept = min(-log10(case_annot$P.Value[case_annot$adj.P.Val < 0.05]))
  ) +
  geom_vline(linetype = "dashed", xintercept = 0.03) +
  geom_vline(linetype = "dashed", xintercept = -0.03) +
  ggtitle("Cases")
# case_annot[case_annot$adj.P.Val <0.05,c("gene","CHR","MAPINFO","logFC","P.Value","UCSC_RefGene_Name")] %>% mutate(UCSC_RefGene_Name = sapply(UCSC_RefGene_Name,function(x)split(x)[1]))

control_annot <- control_res %>%
  rownames_to_column("probe") %>%
  left_join(annot, by = c("probe" = "Name")) %>%
  mutate(gene = gsub(";.*", "", UCSC_RefGene_Name)) %>%
  mutate(gene = ifelse(gene == "", probe, gene))
ggplot(control_annot, aes(logFC, -log10(P.Value))) +
  geom_point() +
  geom_text_repel(
    data = subset(control_annot, adj.P.Val < 0.05 & abs(logFC) > 0.03 & !probe %in% prs_highlight$probe),
    color = "dodgerblue",
    mapping = aes(logFC, -log10(P.Value), label = gene)
  ) +
  geom_point(
    data = subset(control_annot, adj.P.Val < 0.05 & abs(logFC) > 0.03 & !probe %in% prs_highlight$probe),
    color = "red",
    mapping = aes(logFC, -log10(P.Value))
  ) +
  geom_text_repel(
    data = control_annot %>% filter(probe %in% prs_highlight$probe), # subset(control_annot,adj.P.Val < 0.05& abs(logFC) > 0.03),
    color = "dodgerblue",
    mapping = aes(logFC, -log10(P.Value), label = gene)
  ) +
  geom_point(
    data = control_annot %>% filter(probe %in% prs_highlight$probe), # subset(control_annot,adj.P.Val < 0.05 & abs(logFC) > 0.03),
    color = "green",
    mapping = aes(logFC, -log10(P.Value))
  ) +
  labs(x = quote(Delta ~ beta ~ Methylation)) +
  theme_minimal() +
  geom_hline(linetype = "dashed", yintercept = min(-log10(control_annot$P.Value[control_annot$adj.P.Val < 0.05]))) +
  geom_vline(linetype = "dashed", xintercept = 0.03) +
  geom_vline(linetype = "dashed", xintercept = -0.03) +
  ggtitle("Controls")

# control_annot[control_annot$adj.P.Val <0.05,c("gene","CHR","MAPINFO","logFC","P.Value","UCSC_RefGene_Name")] %>% mutate(UCSC_RefGene_Name = sapply(UCSC_RefGene_Name,function(x)split(x)[1]))


ggplot(pd_annot, aes(logFC, -log10(P.Value))) +
  geom_point() +
  geom_text_repel(
    data = pd_annot %>% filter(probe %in% prs_highlight$probe), # subset(pd_annot,adj.P.Val < 0.05& abs(logFC) > 0.03),
    color = "dodgerblue",
    mapping = aes(logFC, -log10(P.Value), label = gene)
  ) +
  geom_point() +
  geom_text_repel(
    data = pd_annot %>% filter(abs(logFC) > 0.03 & adj.P.Val < 0.05) %>% top_n(25, adj.P.Val), # subset(pd_annot,adj.P.Val < 0.05& abs(logFC) > 0.03),
    color = "dodgerblue",
    mapping = aes(logFC, -log10(P.Value), label = gene)
  ) +
  geom_point(
    data = pd_annot %>% filter(probe %in% prs_highlight$probe), # subset(pd_annot,adj.P.Val < 0.05 & abs(logFC) > 0.03),
    color = "green",
    mapping = aes(logFC, -log10(P.Value))
  ) +
  geom_point(
    data = subset(pd_annot, adj.P.Val < 0.05 & abs(logFC) > 0.03 & !probe %in% prs_highlight$probe),
    color = "red",
    mapping = aes(logFC, -log10(P.Value))
  ) +
  labs(x = quote(Delta ~ beta ~ Methylation)) +
  theme_minimal() +
  geom_hline(
    linetype = "dashed",
    yintercept = min(-log10(pd_annot$P.Value[pd_annot$adj.P.Val < 0.05]))
  ) +
  geom_vline(linetype = "dashed", xintercept = 0.03) +
  geom_vline(linetype = "dashed", xintercept = -0.03) +
  ggtitle("PD Diagnosis")
```

```{r}
library(qqman)
qqman::qq(pd_res$P.Value)
manhattan(prs_annot %>% mutate(P = P.Value, SNP = gene, CHR = as.numeric(recode(CHR, X = "23", Y = "24", MT = "26")), BP = MAPINFO), chrlabs = c(1:22, "X", "Y"), annotateTop = F, annotatePval = 5e-8)
qqman::qq(prs_res$P.Value)
hist(prs_res$P.Value, breaks = 100)
hist(pd_res$P.Value, breaks = 100)
```


#### Permutation test for checking if reason for discrepancy between cases and controls is sample sized
```{r, eval=FALSE}
library(parallel)
num_control <- sum(control_select)
num_cases <- sum(case_select)

run_at_sample_size <- function(size) {
  to_select <- sample((1:length(na_row))[na_row], size)
  design_prs <- model.matrix(~ 1 + SCORE1_AVG + age + plate + sex + V3 + V4 + V5 + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = reg_all_data[to_select, ])
  prs_fit <- lmFit(betas_sub[, to_select], design_prs)
  prs_fit <- eBayes(prs_fit)
  topTable(prs_fit, coef = 2, number = Inf)
}


run_at_sample_size_control <- function(size) {
  to_select <- sample((1:length(na_row))[control_select], size)
  design_prs <- model.matrix(~ 1 + SCORE1_AVG + age + plate + sex + V3 + V4 + V5 + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = reg_all_data[to_select, ])
  prs_fit <- lmFit(betas_sub[, to_select], design_prs)
  prs_fit <- eBayes(prs_fit)
  topTable(prs_fit, coef = 2, number = Inf)
}

res_control_size <- mclapply(1:10, function(x) run_at_sample_size(num_control), mc.cores = 10)
res_case_size <- mclapply(1:10, function(x) run_at_sample_size(num_cases), mc.cores = 10)

res_case_size_control <- mclapply(1:10, function(x) run_at_sample_size_control(num_cases), mc.cores = 10)

control_hits <- unlist(lapply(res_control_size, function(x) nrow(subset(x, adj.P.Val < 0.05 & abs(logFC) > 0.03))))

case_hits <- unlist(lapply(res_case_size, function(x) nrow(subset(x, adj.P.Val < 0.05 & abs(logFC) > 0.03))))
case_size_control_hits <- unlist(lapply(res_case_size_control, function(x) nrow(subset(x, adj.P.Val < 0.05 & abs(logFC) > 0.03))))
df <- data.frame(hits = c(mean(case_hits), mean(control_hits)), sample_size = c("Sampled to |Cases|", "Sampled to |Controls|"), dev = c(sd(case_hits), sd(control_hits)))
ggplot(df, aes(x = sample_size, y = hits, ymin = hits - dev, ymax = hits + dev)) +
  geom_bar(stat = "identity") +
  geom_errorbar(width = 0.25) +
  theme_minimal() +
  labs(x = "", y = "Number of hits")
t.test(case_hits, control_hits)
```



#### What are these PRS hits? (EWAS and DMR tables)

```{r}
pd_res %>%
  rownames_to_column("probe") %>%
  filter(P.Value < (0.05 / n())) # , abs(logFC) > 0.1)
write_results <- function(prs_res, prs_dmr, label) {
  prs_res %>%
    rownames_to_column("probe") %>%
    filter(adj.P.Val < 0.25) %>%
    left_join(annot %>% select(Name, CHR, UCSC_RefGene_Name, Relation_to_UCSC_CpG_Island, MAPINFO), by = c("probe" = "Name")) %>%
    mutate(gene = gsub(";.*", "", UCSC_RefGene_Name)) %>%
    mutate(gene = ifelse(gene == "", probe, gene)) %>%
    write.csv(glue("{label}prs_res_fdr025.csv"), row.names = F, quote = F)

  rbindlist(lapply(prs_dmr$coord, function(dmr) as.data.frame(get_dmr_res(dmr, prs_res)))) %>%
    left_join(prs_dmr, by = c("DMR" = "coord")) %>%
    write.csv(glue("{label}PRS_DMR_CPGs.csv"), row.names = F, quote = F)
}
write_results(prs_res, prs_dmr, "cross-sex_")
write_results(male_res, male_prs_dmr, "male_")
write_results(female_res, female_prs_dmr, "female_")
```
## Interaction effects
### PRSxAge interaction
```{r}
design_age <- model.matrix(~ 1 + SCORE1_AVG * age + sex + plate + V3 + V4 + V5 + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = reg_all_data[na_row, ])
age_fit <- lmFit(betas_sub[, na_row], design_age)
age_fit <- eBayes(age_fit)
age_res <- topTable(age_fit, number = Inf, coef = 19)
age_main_res <- topTable(age_fit, number = Inf, coef = c(2, 19))
```
```{r}
ggplot(age_res, aes(logFC, -log10(P.Value))) +
  geom_point() +
  geom_point(
    data = subset(age_res, adj.P.Val < 0.05 & abs(logFC) > 0.03),
    color = "green",
    mapping = aes(logFC, -log10(P.Value))
  ) +
  labs(x = quote(Delta ~ beta ~ Methylation)) +
  theme_minimal() +
  geom_hline(
    linetype = "dashed",
    yintercept = min(-log10(age_res$P.Value[age_res$adj.P.Val < 0.05]))
  ) +
  geom_vline(linetype = "dashed", xintercept = 0.03) +
  geom_vline(linetype = "dashed", xintercept = -0.03)
highlight <- age_res
highlight$adj.P.Val <- age_main_res[rownames(age_res), ]$adj.P.Val
highlight$P.Value <- age_main_res[rownames(age_res), ]$P.Value
ggplot(highlight, aes(logFC, -log10(age_main_res[rownames(age_res), ]$P.Value))) +
  geom_point() +
  geom_point(
    data = subset(highlight, adj.P.Val < 0.05 & abs(logFC) > 0.03),
    color = "green",
    mapping = aes(logFC, -log10(P.Value))
  ) +
  labs(x = quote(Delta ~ beta ~ Methylation)) +
  theme_minimal() +
  geom_hline(
    linetype = "dashed",
    yintercept = min(-log10(age_main_res$P.Value[age_main_res$adj.P.Val < 0.05]))
  ) +
  geom_vline(linetype = "dashed", xintercept = 0.03) +
  geom_vline(linetype = "dashed", xintercept = -0.03)
```
### PRSxSex interaction

```{r}
design_sex <- model.matrix(~ 1 + SCORE1_AVG * sex + age + plate + V3 + V4 + V5 + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = reg_all_data[na_row, ])
sex_fit <- lmFit(betas_sub[, na_row], design_sex)
sex_fit <- eBayes(sex_fit)
sex_res <- topTable(sex_fit, number = Inf, coef = 19)
sex_main_res <- topTable(sex_fit, number = Inf, coef = c(2, 19))
```
```{r}
ggplot(sex_res, aes(logFC, -log10(P.Value))) +
  geom_point() +
  geom_point(
    data = subset(sex_res, adj.P.Val < 0.05 & abs(logFC) > 0.03),
    color = "green",
    mapping = aes(logFC, -log10(P.Value))
  ) +
  labs(x = quote(Delta ~ beta ~ Methylation)) +
  theme_minimal() +
  geom_hline(
    linetype = "dashed",
    yintercept = min(-log10(sex_res$P.Value[sex_res$adj.P.Val < 0.05]))
  ) +
  geom_vline(linetype = "dashed", xintercept = 0.03) +
  geom_vline(linetype = "dashed", xintercept = -0.03)
highlight <- sex_res
highlight$adj.P.Val <- sex_main_res[rownames(sex_res), ]$adj.P.Val
highlight$P.Value <- sex_main_res[rownames(sex_res), ]$P.Value
ggplot(highlight, aes(logFC, -log10(sex_main_res[rownames(sex_res), ]$P.Value))) +
  geom_point() +
  geom_point(
    data = subset(highlight, adj.P.Val < 0.05 & abs(logFC) > 0.03),
    color = "green",
    mapping = aes(logFC, -log10(P.Value))
  ) +
  labs(x = quote(Delta ~ beta ~ Methylation)) +
  theme_minimal() +
  geom_hline(
    linetype = "dashed",
    yintercept = min(-log10(sex_main_res$P.Value[sex_main_res$adj.P.Val < 0.05]))
  ) +
  geom_vline(linetype = "dashed", xintercept = 0.03) +
  geom_vline(linetype = "dashed", xintercept = -0.03)
```
#### Are these probes different than those with genetic main effect
```{r}
prs_hits <- prs_res %>%
  rownames_to_column("probe") %>%
  filter(P.Value < (0.05 / n()))
sex_main_res %>%
  rownames_to_column("probe") %>%
  filter(P.Value < (0.05 / n()), !probe %in% prs_hits$probe)
age_main_res %>%
  rownames_to_column("probe") %>%
  filter(P.Value < (0.05 / n()), !probe %in% prs_hits$probe)
ggplot(reg_all_data, aes(SCORE1_AVG, betas_sub[rownames(sex_main_res[which.min(sex_main_res$adj.P.Val), ]), ], color = sex)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(y = glue("{rownames(sex_main_res[which.min(sex_main_res$P.Value),])} Methylation"), x = "PD PRS")
```

### PRSxExposure Interaction
Loading in multiple imputation data
```{r}
library(mice)
pest_missing <- read.csv("/home1/NEURO/SHARE_DECIPHER/TERRE_pesticides/pesticides.csv")
# pest_missing <- pest_missing[pest_missing$num %in%colnames(betas_sub), ]
pest_imputed <- read.csv("/home1/NEURO/SHARE_DECIPHER/TERRE_pesticides/pesticides_imputed.csv")
# pest_imputed <- pest_imputed[pest_imputed$num %in%colnames(betas_sub), ]
pre_mids <- rbind(cbind(data.frame(X_imputation_ = 0), pest_missing)[, colnames(pest_imputed)], pest_imputed)
pre_mids <- pre_mids[pre_mids$num %in% colnames(betas_sub), ]
pest_mids <- as.mids(pre_mids, .imp = "X_imputation_", .id = "num")
## @TODO RE-RUN ANALYSES AND AGGREGATE WITH RUBIN'S METHOD WITH mice PACKAGE
betas_pest <- betas_sub[, rownames(pest_mids$data)]
exposures <- colnames(pest_mids$data)[colMeans(pest_mids$data, na.rm = T) > 0.025]
```
### PRS exposure interaction with Random Forest Imputation
```{r}
library(missForest)
to_impute <- pest_missing[, -c(1)] %>% mutate_all(~ factor(.))
pest_forest <- missForest(to_impute)
pest_forest$OOBerror
pest_forest_imputed <- cbind(num = pest_missing$num, pest_forest$ximp %>% mutate_all(~ as.numeric(as.character(.))))
reg_w_all_pesticide <- reg_all_data %>%
  left_join(pest_forest_imputed, by = c("patient" = "num")) %>%
  filter(!is.na(ins))

reg_pre_impute <- reg_all_data %>%
  select(patient, SCORE1_AVG, V3, V4, V5, age, men, MMS) %>%
  left_join(pest_missing %>% mutate_all(~ factor(.)), by = c("patient" = "num"))
pest_forest_w_vars <- missForest(reg_pre_impute[, -c(1:5)])
reg_all_vars <- cbind(reg_pre_impute[, 1:8], pest_forest_w_vars$ximp[, -c(1:3)] %>% mutate_all(~ as.numeric(as.character(.)))) %>% filter(patient %in% reg_w_all_pesticide$patient)
pest_forest_w_vars$OOBerror
```

Run regressions:
```{r}
betas_pest <- betas_sub[, as.character(reg_w_all_pesticide$patient)]
all(reg_w_all_pesticide$patient == colnames(betas_pest))
exposures <- colnames(pest_forest$ximp)[colMeans(reg_w_all_pesticide[, .SD, .SDcols = colnames(pest_forest$ximp)]) > 0.025]
# exposures_vars <- colnames(pest_forest_w_vars$ximp[,-c(1:3)])[colMeans(reg_all_vars[,.SD,.SDcols=colnames(pest_forest_w_vars$ximp[,-c(1:3)])]) > 0.04]
res <- parallel::mclapply(
  exposures,
  function(x) {
    design <- substitute(model.matrix(~ E * SCORE1_AVG + sex + age + V3 + V4 + V5, reg_w_all_pesticide), list(E = as.name(x)))
    fit <- lmFit(betas_pest, eval(design))
    eb <- eBayes(fit)
    topTable(eb, coef = 9, number = Inf)
  },
  mc.cores = 16
)
names(res) <- as.character(exposures)

# res_vars <- parallel::mclapply(
#     exposures_vars,
#     function(x){
#       design <- substitute(model.matrix(~E*SCORE1_AVG + men + age + V3 + V4 + V5,reg_all_vars),list(E=as.name(x)))
#       fit <- lmFit(betas_pest,eval(design))
#       eb <- eBayes(fit)
#       topTable(eb,coef=9,number = Inf)
#     },
#     mc.cores = 16)
# names(res_vars) <- as.character(exposures_vars)

all_pest_prs_df <- rbindlist(lapply(res, function(df) rownames_to_column(df, var = "probe")), idcol = "exposure") %>%
  filter(adj.P.Val < 0.25) %>%
  left_join(annot, by = c("probe" = "Name"))
(prs_int_hits <- all_pest_prs_df %>% arrange(P.Value) %>% filter(adj.P.Val < 0.05) %>% select(exposure, probe, logFC, P.Value, adj.P.Val, UCSC_RefGene_Name))
# all_pest_prs_vars_df <- rbindlist(lapply(res_vars,function(df) rownames_to_column(df,var = "probe")),idcol="exposure") %>% filter(adj.P.Val < 0.25) %>% left_join(annot,by=c("probe"="Name"))
# all_pest_prs_vars_df %>% arrange(P.Value)
prs_fong_s_cu <- rbindlist(lapply(res, function(df) rownames_to_column(df, var = "probe")), idcol = "exposure") %>%
  filter(exposure %in% c("fong", "s", "cu")) %>%
  left_join(annot, by = c("probe" = "Name"))
print(prs_int_hits)
```
```{r}
to_plot <- prs_fong_s_cu %>% mutate(exposure = recode(exposure, cu = "Copper", fong = "Fungicides", s = "Sulphur"))
ggplot(to_plot, aes(logFC, -log10(P.Value), color = exposure)) +
  geom_point() +
  geom_text_repel(data = subset(to_plot, adj.P.Val < 0.05), mapping = aes(logFC, -log10(P.Value), label = UCSC_RefGene_Name)) +
  labs(color = "Exposure", x = quote(Delta ~ beta ~ Methylation)) +
  theme_minimal()
```

```{r}
methy_tmp <- betas_pest["cg13975855", ]
ggplot(reg_w_all_pesticide, aes(SCORE1_AVG, methy_tmp, color = ifelse(fong == 1 | cu == 1 | s == 1, "Exposed", "Not Exposed"))) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  facet_grid(~PD) +
  labs(x = "PRS for PD", color = "Copper, sulphur, or fungicide", y = quote(beta ~ Methylation ~ of ~ cg13975855)) +
  theme_minimal()
reg_w_all_pesticide <- reg_w_all_pesticide %>% mutate(combined = ifelse(fong == 1 | cu == 1 | s == 1, "Exposed", "Not Exposed"))
ggplot(reg_w_all_pesticide, aes(SCORE1_AVG, methy_tmp)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(data = reg_w_all_pesticide, aes(color = PD)) +
  facet_grid(~combined) +
  labs(x = "Polygenic risk score for PD", y = quote(beta ~ Methylation ~ of ~ cg13975855), color = "PD status") +
  theme_minimal()
ggplot(reg_w_all_pesticide, aes(SCORE1_AVG, methy_tmp, color = cu == 1)) +
  geom_point() +
  geom_smooth(method = "lm")
ggplot(reg_w_all_pesticide, aes(SCORE1_AVG, methy_tmp, color = s == 1)) +
  geom_point() +
  geom_smooth(method = "lm")
```

# CASE and CONTROL only analysis
## Pre-check that PRS is independent of exposure

```{r}
prs_vs_pest <- lapply(exposures, function(x) t.test(reg_w_all_pesticide$SCORE1_AVG ~ unlist(reg_w_all_pesticide[, ..x]))$p.value)
names(prs_vs_pest) <- exposures
prs_vs_pest %>%
  stack() %>%
  arrange(values)
exposures_ind <- names(prs_vs_pest[prs_vs_pest > 0.05])
```
Checking correlation between exposures:
```{r}
library(pheatmap)
cor(reg_w_all_pesticide[, ..exposures_ind]) %>%
  pheatmap(main = "Correlation between exposures")
# melt() %>%
# ggplot(aes(Var1,Var2,fill=value)) + geom_tile() + scale_fill_viridis_c()+
# theme(axis.text.x=element_text(angle=90,hjust=1,vjust=1))
cor(reg_w_all_pesticide[PD == "CASE", ..exposures_ind]) %>%
  pheatmap(main = "Correlation between exposures cases only")
cor(reg_w_all_pesticide[PD == "CONTROL", ..exposures_ind]) %>%
  pheatmap(main = "Correlation between exposures controls only")
```

## TERRE CASE ONLY ANALYSIS
```{r}
betas_pest_case <- betas_pest[, reg_w_all_pesticide$PD == "CASE"]
reg_w_all_pesticide_case <- reg_w_all_pesticide %>% filter(PD == "CASE")
all(reg_w_all_pesticide$patient == colnames(betas_pest))
exposures <- colnames(pest_forest$ximp)[colMeans(reg_w_all_pesticide_case[, .SD, .SDcols = colnames(pest_forest$ximp)]) > 0.025 & colnames(pest_forest$ximp) %in% exposures_ind]
res_case <- parallel::mclapply(
  exposures,
  function(x) {
    design <- substitute(model.matrix(~ E * SCORE1_AVG + sex + age + V3 + V4 + V5, reg_w_all_pesticide_case), list(E = as.name(x)))
    fit <- lmFit(betas_pest_case, eval(design))
    eb <- eBayes(fit)
    topTable(eb, coef = 9, number = Inf)
  },
  mc.cores = 16
)
names(res_case) <- as.character(exposures)
all_pest_prs_case_df <- rbindlist(lapply(res_case, function(df) rownames_to_column(df, var = "probe")), idcol = "exposure") %>%
  filter(adj.P.Val < 0.05) %>%
  left_join(annot, by = c("probe" = "Name"))
all_pest_prs_case_df %>%
  arrange(P.Value) %>%
  select(exposure, probe, logFC, AveExpr, t, P.Value, adj.P.Val, B, CHR, MAPINFO, UCSC_RefGene_Name) %>%
  View()
```
```{r}
to_plot <- rbindlist(lapply(res_case, function(df) rownames_to_column(df, var = "probe")), idcol = "exposure") %>%
  filter(exposure %in% unique(all_pest_prs_case_df$exposure[order(all_pest_prs_case_df$adj.P.Val)])[1:5]) %>%
  left_join(annot, by = c("probe" = "Name")) %>%
  mutate(gene = gsub(";.*", "", UCSC_RefGene_Name))
ggplot(to_plot, aes(logFC, -log10(P.Value), color = exposure)) +
  geom_point() +
  geom_text_repel(data = subset(to_plot, adj.P.Val < 0.05), mapping = aes(logFC, -log10(P.Value), label = gene)) +
  labs(color = "Exposure", x = quote(Delta ~ beta ~ Methylation)) +
  theme_minimal()
```


# TERRE control ONLY ANALYSIS
```{r}
betas_pest_control <- betas_pest[, reg_w_all_pesticide$PD == "CONTROL"]
reg_w_all_pesticide_control <- reg_w_all_pesticide %>% filter(PD == "CONTROL")
all(reg_w_all_pesticide$patient == colnames(betas_pest))
exposures <- colnames(pest_forest$ximp)[colMeans(reg_w_all_pesticide_control[, .SD, .SDcols = colnames(pest_forest$ximp)]) > 0.025 & colnames(pest_forest$ximp) %in% exposures_ind]
res_control <- parallel::mclapply(
  exposures,
  function(x) {
    design <- substitute(model.matrix(~ E * SCORE1_AVG + sex + age + V3 + V4 + V5, reg_w_all_pesticide_control), list(E = as.name(x)))
    fit <- lmFit(betas_pest_control, eval(design))
    eb <- eBayes(fit)
    topTable(eb, coef = 9, number = Inf)
  },
  mc.cores = 16
)
names(res_control) <- as.character(exposures)
all_pest_prs_control_df <- rbindlist(lapply(res_control, function(df) rownames_to_column(df, var = "probe")), idcol = "exposure") %>%
  filter(adj.P.Val < 0.05) %>%
  left_join(annot, by = c("probe" = "Name"))
all_pest_prs_control_df %>%
  arrange(P.Value) %>%
  select(exposure, probe, logFC, AveExpr, t, P.Value, adj.P.Val, B, CHR, MAPINFO, UCSC_RefGene_Name) %>%
  View()
```
```{r}
to_plot <- rbindlist(lapply(res_control, function(df) rownames_to_column(df, var = "probe")), idcol = "exposure") %>%
  filter(exposure %in% unique(all_pest_prs_control_df$exposure[order(all_pest_prs_control_df$adj.P.Val)])[1:5]) %>%
  left_join(annot, by = c("probe" = "Name")) %>%
  mutate(gene = gsub(";.*", "", UCSC_RefGene_Name))
ggplot(to_plot, aes(logFC, -log10(P.Value), color = exposure)) +
  geom_point() +
  geom_text_repel(data = subset(to_plot, adj.P.Val < 0.05), mapping = aes(logFC, -log10(P.Value), label = gene)) +
  labs(color = "Exposure", x = quote(Delta ~ beta ~ Methylation)) +
  theme_minimal()
```
# DIGPD PRS Analysis
Check SNPs in set of variants:

```{r}
bim <- fread("~/genotype_qc/DIGPD_merged_post_imputation_QC/raw_data.imputed.r2_90.hard_call.maf_mismatch.bim", header = F)
sum(risk_scores$SNP %in% bim$V2)
```

Generate PRS using `plink2 --score`:
```{bash, eval=FALSE}
~/plink2 --bfile ~/genotype_qc/DIGPD_merged_post_imputation_QC/raw_data.imputed.r2_90.hard_call.maf_mismatch\
  --score pd_prs_file.txt list-variants\
  --out digpd_prs

~/plink2 --bfile ~/genotype_qc/DIGPD_case_QC/all_imputed_r2_30_rsid\
  --score pd_prs_file.txt list-variants\
  --out digpd_control_prs

~/plink2 --bfile ~/genotype_qc/DIGPD_control_QC/all_imputed_r2_30_rsid\
  --score pd_prs_file.txt list-variants\
  --out digpd_case_prs


```
## Loading in and preparing data
```{r}
digpd_merged_fam <- rbind(fread("~/genotype_qc/DIGPD_case_pre_pca.matched.fam"), fread("~/genotype_qc/DIGPD_control_pre_pca.matched.fam"))
digpd_prs <- fread("digpd_prs.sscore")
digpd_prs$sex <- ifelse(digpd_merged_fam$V5 == 1, "MALE", "FEMALE")
digpd_prs$PD <- ifelse(digpd_merged_fam$V6 == 2, "CASE", "CONTROL")
digpd_control_prs <- fread("digpd_control_prs.sscore")
digpd_case_prs <- fread("digpd_case_prs.sscore")
strat_prs <- rbind(digpd_control_prs, digpd_case_prs)
digpd_pcs <- fread("~/genotype_qc/DIGPD_merged_post_imputation_QC/raw_data.preimpute.pcs")
digpd_metadata <- fread("/home1/NEURO/SHARE_DECIPHER/digpd_meta_master.csv") %>%
  select(-PD) %>%
  mutate(IID = gsub("PAL_[0-9]*_", "", IID))
all_digpd_data <- digpd_prs %>%
  left_join(digpd_metadata, by = c("IID")) %>%
  left_join(digpd_pcs, by = c("IID"))
```

### PRS score testing

```{r}
ggplot(all_digpd_data, aes(PD, SCORE1_AVG)) +
  geom_boxplot() +
  geom_jitter(height = 0, width = 0.1) +
  labs(x = "Parkinson's Diagnosis", y = "Polygenic Risk Score") +
  ggtitle(bquote("Nagelkerke Pseudo-" * R^2 * ":0.0578120"))
t.test(all_digpd_data$SCORE1_AVG[all_digpd_data$PD == "CASE"], all_digpd_data$SCORE1_AVG[all_digpd_data$PD == "CONTROL"], alternative = "greater")
```

### PRS DIGPD PD associations
```{r}
all_digpd_data$PD_fixed <- ifelse(all_digpd_data$PD == "CASE", 1, 0)
fit1 <- glm(PD_fixed ~ SCORE1_AVG + sex + PC1 + PC2 + PC3, data = all_digpd_data, family = "binomial", control = list(maxit = 50))
summary(fit1)
fit1_null <- glm(PD_fixed ~ sex + PC1 + PC2 + PC3, data = all_digpd_data, family = "binomial", control = list(maxit = 50))
summary(fit1_null)
nagelkerke(fit1, null = fit1_null)
lin1 <- lm(PD_fixed ~ SCORE1_AVG + sex + PC1 + PC2 + PC3, data = all_digpd_data)
lin2 <- lm(PD_fixed ~ sex + PC1 + PC2 + PC3, data = all_digpd_data)
R2 <- (deviance(lin2) - deviance(lin1)) / deviance(lin2)
h2 <- pd_h2(R2, sum(all_digpd_data$PD_fixed) / nrow(all_digpd_data), 0.01)
fit2 <- glm(PD_fixed ~ SCORE1_AVG * sex + PC1 + PC2 + PC3, data = all_digpd_data, family = "binomial")
summary(fit2)
fit3 <- glm(PD_fixed ~ SCORE1_AVG * smoking + sex + PC1 + PC2 + PC3, data = all_digpd_data, family = "binomial")
summary(fit3)
fit4 <- glm(PD_fixed ~ SCORE1_AVG * pesticides + sex + PC1 + PC2 + PC3, data = all_digpd_data, family = "binomial")
summary(fit4)
```
## DIGPD PRS methylation associations
```{r}
load("/home1/NEURO/SHARE_DECIPHER/processed_DNAm_data/DIGPD_processed_2021/DIGPD_betas_combat.RData")
betas_sub <- betas_combat
all_digpd_data$patient <- paste0(all_digpd_data$patient, ".visit", all_digpd_data$visite)
digpd_reg_data <- all_digpd_data %>%
  filter(patient %in% colnames(betas_sub)) %>%
  mutate(plate = factor(plate))
betas_digpd <- betas_sub[, digpd_reg_data$patient]
```
### including covariates

```{r}
pcs <- prcomp(t(betas_digpd))
```
```{r}
library(pheatmap)
tmp_design <- model.matrix(~ 0 + ., data = digpd_reg_data[, c("race1", "plate", "birth_cohort")] %>% mutate(race1 = factor(race1), birth_cohort = birth_cohort - min(birth_cohort)))
pheatmap(
  cor(pcs$x[, 1:20],
    cbind(tmp_design, digpd_reg_data[, c("men", "PD", "levodopa", "Age", "pesticides", "durevol", "smoking", "bmi")]) %>% mutate(PD = as.numeric(PD == "CASE")),
    use = "pairwise.complete.obs"
  ),
  cluster_rows = F
)
```

### digpd PRS association
```{r}
design_prs <- model.matrix(~ 1 + SCORE1_AVG + Age + men + PC1 + PC2 + PC3 + birth_cohort, data = digpd_reg_data %>% mutate(race1 = factor(race1), birth_cohort = birth_cohort - min(birth_cohort), PD = as.numeric(PD == "CASE")))
prs_fit <- lmFit(betas_digpd, design_prs)
prs_fit <- eBayes(prs_fit)
design_pd <- model.matrix(~ 1 + PD + Age + men + plate + birth_cohort, data = digpd_reg_data %>% mutate(race1 = factor(race1), birth_cohort = birth_cohort - min(birth_cohort), PD = as.numeric(PD == "CASE")))
pd_fit <- lmFit(betas_digpd, design_pd)
pd_fit <- eBayes(pd_fit)
pd_res <- topTable(pd_fit, coef = 2, number = Inf)
prs_res <- topTable(prs_fit, coef = 2, number = Inf)
```


```{r}
library(ggrepel)
annot <- fread("~/MethylationEPIC_v-1-0_B4.csv", skip = 7)
pd_annot_digpd <- pd_res %>%
  rownames_to_column("probe") %>%
  left_join(annot, by = c("probe" = "Name")) %>%
  mutate(gene = gsub(";.*", "", UCSC_RefGene_Name)) %>%
  mutate(gene = ifelse(gene == "", probe, gene))
prs_annot_digpd <- prs_res %>%
  rownames_to_column("probe") %>%
  left_join(annot, by = c("probe" = "Name")) %>%
  mutate(gene = gsub(";.*", "", UCSC_RefGene_Name)) %>%
  mutate(gene = ifelse(gene == "", probe, gene))

ggplot(pd_annot_digpd, aes(logFC, -log10(P.Value))) +
  geom_point() +
  geom_text_repel(data = pd_annot_digpd %>% filter(abs(logFC) > 0.03 & adj.P.Val < 0.05) %>% top_n(25, -adj.P.Val), mapping = aes(logFC, -log10(P.Value), label = gene)) +
  geom_point(data = pd_annot_digpd %>% filter(abs(logFC) > 0.03 & adj.P.Val < 0.05), color = "red", mapping = aes(logFC, -log10(P.Value))) +
  labs(x = quote(Delta ~ beta ~ Methylation)) +
  theme_minimal() +
  geom_hline(linetype = "dashed", yintercept = min(-log10(pd_annot_digpd$P.Value[pd_annot_digpd$adj.P.Val < 0.05]))) +
  geom_vline(linetype = "dashed", xintercept = 0.03) +
  geom_vline(linetype = "dashed", xintercept = -0.03)

ggplot(prs_annot_digpd, aes(logFC, -log10(P.Value))) +
  geom_point() +
  geom_text_repel(data = subset(prs_annot_digpd, adj.P.Val < 0.05 & abs(logFC) > 0.03), mapping = aes(logFC, -log10(P.Value), label = gene)) +
  geom_point(data = subset(prs_annot_digpd, adj.P.Val < 0.05 & abs(logFC) > 0.03), color = "red", mapping = aes(logFC, -log10(P.Value))) +
  labs(x = quote(Delta ~ beta ~ Methylation)) +
  theme_minimal() +
  geom_hline(linetype = "dashed", yintercept = min(-log10(prs_annot_digpd$P.Value[prs_annot_digpd$adj.P.Val < 0.05]))) +
  geom_vline(linetype = "dashed", xintercept = 0.03) +
  geom_vline(linetype = "dashed", xintercept = -0.03)
```
### checking reason for difference in hits between DIGPD and Terre
Suspicious inflation here:
```{r}
qq(pd_annot_digpd$P.Value, main = "DIGPD Parkinson's EWAS")
qq(pd_annot$P.Value, main = "Terre Parkinson's EWAS")
qq(prs_annot_digpd$P.Value, main = "DIGPD Parkinson's PRS EWAS")
qq(prs_annot$P.Value, main = "Terre Parkinson's PRS EWAS")
```

```{r}
library(qqman)
manhattan(prs_annot_digpd %>% mutate(P = P.Value, SNP = gene, CHR = as.numeric(recode(CHR, X = "23", Y = "24", MT = "26")), BP = MAPINFO), chrlabs = c(1:22, "X", "Y"), annotateTop = F, annotatePval = 5e-8)
```
```{r}
qq(pd_annot_digpd$P.Value, main = "DIGPD Parkinson's EWAS")
```
