---
title: "Post mQTL analysis Terre"
output: html_notebook
---

```{r setup, include=FALSE}
library(coloc)
library(data.table)
library(parallel)
library(qqman)
library(qvalue)
library(readxl)
library(Gviz)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```
## Replication of Cross-sex mQTL
```{r}
library(dbplyr)
library(DBI)
ntests <- 126904049
file.exists("terre_data/cis_all_impute_mQTL_results_9_methy_PC.db")
con_terre <- dbConnect(RSQLite::SQLite(), "terre_data/cis_all_impute_mQTL_results_9_methy_PC.db")

mQTL_results <- tbl(con_terre, "datatable")
con_digpd <- dbConnect(RSQLite::SQLite(), "digpd_data/cis_all_impute_mQTL_results_9_methy_PC.db")
ntests_digpd <- 100363748
mQTL_results_digpd <- tbl(con_digpd, "datatable")
```
```{r}
mQTL_results %>%
  filter(FDR < 0.05) %>%
  summarize(total = n(), cpgs = n_distinct(gene))
mQTL_results %>%
  filter(`p-value` < (0.05 / ntests)) %>%
  summarize(total = n(), cpgs = n_distinct(gene))

mQTL_results_digpd %>%
  filter(FDR < 0.05) %>%
  summarize(total = n(), cpgs = n_distinct(gene))
mQTL_results_digpd %>%
  filter(`p-value` < (0.05 / ntests)) %>%
  summarize(total = n(), cpgs = n_distinct(gene))

library(dbplot)
mQTL_results %>% dbplot_histogram(`p-value`)
# qq(mQTL_results$`p-value`)
```

### Checking replication
```{r}
to_check_replication <- mQTL_results_digpd %>%
  inner_join(mQTL_results %>% filter(FDR < 0.05), by = c("SNP", "gene"), copy = TRUE) %>%
  collect()
```

$\pi_1$ statistics
```{r}
to_check_replication <- as.data.table(to_check_replication)

pi0res <- pi0est(na.omit(as.numeric(to_check_replication$`p-value.x`)), lambda = seq(0, 0.245, 0.005), pi0.method = "bootstrap")
1 - pi0res$pi0
plot(pi0res$pi0.lambda, pi0res$lambda, xlab = bquote(pi[0]), ylab = bquote(lambda))
abline(v = pi0res$pi0)
qvalres <- qvalue_truncp(na.omit(as.numeric(to_check_replication$`p-value.x`)))
1 - qvalres$pi0
abline(v = qvalres$pi0)
pi0est(as.numeric(to_check_replication$`p-value.x`) / 0.25)
```

## Colocalization with parkinson's loci
```{r colocalization, eval=FALSE}
probe_pos <- fread("terre_data/probe_pos.txt")
snp_pos <- fread("terre_data/snp_pos.txt", key = "SNP")
methy <- fread("~/prs_ewas_integration/cis_mQTL_analyses/terre_data/methylation_combat.txt", key = "cpg")
sum_stats <- fread("~/nallsEtAl2019_excluding23andMe_allVariants.tab.gz", key = "SNP")
risk_scores <- read_xlsx("~/pd_risk_score/Supplementary materials/Table S2. Detailed summary statistics on all nominated risk variants, known and novel_.xlsx") %>%
  mutate(CHR = as.numeric(CHR), COUNTED = toupper(`Effect allele`), ALT = toupper(`Other allele`), POS = BP) %>%
  dplyr::filter(!is.na(CHR))
mQTL_results <- fread("terre_data/cis_all_impute_mQTL_results_9_methy_PC.txt.gz", key = "gene")

# Grab CpG sites within 1MB of risk loci
probes_1mb <- function(row) {
  probe_pos[chr == paste0("chr", row$CHR) & s1 >= row$POS - 1e6 & s1 <= row$POS + 1e6]$geneid
}

localize_per_locus <- function(i, mQTL_results = mQTL_results, snp_pos = snp_pos, has_methy = TRUE, N1 = 245, return_summary = TRUE) {
  all_probes <- unlist(probes_1mb(risk_scores[i, ]))
  probe_res <- mclapply(
    1:length(all_probes),
    function(j) {
      snp_set <- mQTL_results[all_probes[j], on = "gene"][!duplicated(SNP)]
      coords <- snp_pos[snp_set$SNP, on = "SNP"][!duplicated(SNP)]
      tmp_stats <- sum_stats[paste0(coords$CHR, ":", coords$POS), on = "SNP"]
      if (has_methy) {
        mqtl_pval <- data.frame(
          beta = snp_set$beta,
          varbeta = snp_set$beta / snp_set$`t-stat`,
          N = N1,
          snp = paste0(coords$CHR, ":", coords$POS),
          sdY = sd(methy[all_probes[j], -c(1), on = "cpg"]),
          type = "quant",
          stringsAsFactors = F
        )
      } else {
        mqtl_pval <- data.frame(
          beta = snp_set$beta,
          N = N1,
          snp = paste0(coords$CHR, ":", coords$POS),
          MAF = tmp_stats$freq,
          pvalues = snp_set$`p-value`,
          type = "quant",
          stringsAsFactors = F
        )
      }
      gwas_pval <- data.frame(
        pvalues = tmp_stats$p,
        N = tmp_stats$N_cases + tmp_stats$N_controls,
        s = tmp_stats$N_cases / (tmp_stats$N_cases + tmp_stats$N_controls),
        snp = tmp_stats$SNP,
        type = "cc",
        MAF = tmp_stats$freq,
        stringsAsFactors = F
      )
      mqtl_pval <- as.list(mqtl_pval[!is.na(gwas_pval$pvalues), ])
      mqtl_pval$N <- unique(mqtl_pval$N)[1]
      mqtl_pval$type <- unique(mqtl_pval$type)[1]
      gwas_pval <- as.list(gwas_pval[!is.na(gwas_pval$pvalues), ])
      gwas_pval$N <- unique(gwas_pval$N)[1]
      gwas_pval$type <- unique(gwas_pval$type)[1]
      print(j)
      if (length(gwas_pval$snp) != 0 | length(mqtl_pval$snp != 0)) {
        res <- coloc.abf(mqtl_pval, gwas_pval)
        if (return_summary) {
          return(cbind(t(data.frame(res$summary)), data.frame(probe = all_probes[j], locus = risk_scores$`Locus Number`[i], locus_snp = risk_scores$SNP[i])))
        } else {
          return(cbind(res$results, data.frame(probe = all_probes[j], locus = risk_scores$`Locus Number`[i], locus_snp = risk_scores$SNP[i])))
        }
      } else {
        return(data.frame())
      }
    },
    mc.cores = 8
  )
  return(rbindlist(probe_res))
}
# system.time(res1 <- rbindlist(mclapply(1:nrow(risk_scores), function(i) localize_per_locus(i, mQTL_results = mQTL_results, snp_pos = snp_pos, has_methy = TRUE, N1 = 245), mc.cores = 4)))
system.time(res1_pval <- rbindlist(mclapply(1:nrow(risk_scores), function(i) localize_per_locus(i, mQTL_results = mQTL_results, snp_pos = snp_pos, has_methy = FALSE, N1 = 245), mc.cores = 4)))

system.time(res1_SNP <- rbindlist(mclapply(1:nrow(risk_scores), function(i) localize_per_locus(i, mQTL_results = mQTL_results, snp_pos = snp_pos, has_methy = FALSE, N1 = 245, return_summary = FALSE), mc.cores = 4)))
fwrite(res1_SNP, "mqtl_pd_snp_colocalization.txt.gz", sep = "\t", row.names = F, quote = F)
fwrite(res1_pval, "mqtl_pd_snp_colocalization.txt", sep = "\t", row.names = F, quote = F)
```
## Sex stratified colocalization

```{r}

```


## Plot out results
```{r}
if (file.exists("mqtl_pd_snp_colocalization.txt")) {
  res1_pval <- fread("mqtl_pd_snp_colocalization.txt")
  probe_pos <- fread("terre_data/probe_pos.txt")
}
res1_pval %>%
  mutate(PP3_4 = PP.H3.abf + PP.H4.abf, PP_4div3 = PP.H4.abf / PP.H3.abf) %>%
  arrange(-PP3_4, -PP_4div3) %>%
  filter(PP3_4 > 0.99, PP_4div3 > 5)
res1_pval %>%
  mutate(PP3_4 = PP.H3.abf + PP.H4.abf, PP_4div3 = PP.H4.abf / PP.H3.abf) %>%
  arrange(-PP3_4, -PP_4div3) %>%
  filter(PP3_4 > 0.99, PP_4div3 > 5) %>%
  distinct(probe, locus)
res1_pval %>%
  mutate(PP3_4 = PP.H3.abf + PP.H4.abf, PP_4div3 = PP.H4.abf / PP.H3.abf) %>%
  arrange(-PP3_4, -PP_4div3) %>%
  filter(PP3_4 > 0.99, PP_4div3 > 5) %>%
  distinct(probe)
res1_pval %>%
  mutate(PP3_4 = PP.H3.abf + PP.H4.abf, PP_4div3 = PP.H4.abf / PP.H3.abf) %>%
  arrange(-PP3_4, -PP_4div3) %>%
  filter(PP3_4 > 0.99, PP_4div3 > 5) %>%
  left_join(probe_pos, by = c("probe" = "geneid")) %>%
  distinct(locus, chr)
```
```{r}
res1_pval %>%
  mutate(PP3_4 = PP.H3.abf + PP.H4.abf, PP_4div3 = PP.H4.abf / PP.H3.abf) %>%
  arrange(-PP3_4, -PP_4div3) %>%
  filter(PP3_4 > 0.99, PP_4div3 > 5) %>%
  write_delim("mqtl_pd_snp_colocalized_blood.txt")

res1_pval %>%
  filter(PP.H4.abf > 0.9) %>%
  write_delim("mqtl_pd_snp_colocalized_blood_PH4.txt")
```

modest evidence of colocalization between parkinson's loci and mQTL in blood of TERRE cohort, let's try this analysis on top mQTL CpGs instead:
```{r}
all_probes <- mQTL_results[`p-value` < 0.05 / (ntests), .(unique(gene))]$V1
probe_res <- mclapply(
  1:length(all_probes),
  function(j) {
    snp_set <- mQTL_results[all_probes[j], on = "gene"][!duplicated(SNP)]
    coords <- snp_pos[snp_set$SNP, on = "SNP"][!duplicated(SNP)]
    mqtl_pval <- data.frame(beta = snp_set$beta, varbeta = snp_set$beta / snp_set$`t-stat`, N = 245, snp = paste0(coords$CHR, ":", coords$POS), sdY = sd(methy[all_probes[j], -c(1), on = "cpg"]), type = "quant", stringsAsFactors = F)
    tmp_stats <- sum_stats[mqtl_pval$snp, on = "SNP"]
    gwas_pval <- data.frame(pvalues = tmp_stats$p, N = tmp_stats$N_cases + tmp_stats$N_controls, s = tmp_stats$N_cases / (tmp_stats$N_cases + tmp_stats$N_controls), snp = tmp_stats$SNP, type = "cc", MAF = tmp_stats$freq, stringsAsFactors = F)
    mqtl_pval <- as.list(mqtl_pval[!is.na(gwas_pval$pvalues), ])
    mqtl_pval$N <- unique(mqtl_pval$N)[1]
    mqtl_pval$type <- unique(mqtl_pval$type)[1]
    gwas_pval <- as.list(gwas_pval[!is.na(gwas_pval$pvalues), ])
    gwas_pval$N <- unique(gwas_pval$N)[1]
    gwas_pval$type <- unique(gwas_pval$type)[1]
    print(j)
    if (length(gwas_pval$snp) != 0 | length(mqtl_pval$snp != 0)) {
      res <- coloc.abf(mqtl_pval, gwas_pval)
      return(cbind(t(data.frame(res$summary)), data.frame(probe = all_probes[j])))
    } else {
      return(data.frame())
    }
  },
  mc.cores = 64
)

probe_res <- rbindlist(probe_res)

fwrite(probe_res, "mqtl_pd_probe_colocalization.txt", sep = "\t", row.names = F, quote = F)
```

```{r}
if (file.exists("mqtl_pd_probe_colocalization.txt")) {
  probe_res <- fread("mqtl_pd_probe_colocalization.txt")
}

probe_res %>%
  mutate(PP3_4 = PP.H3.abf + PP.H4.abf, PP_4div3 = PP.H4.abf / PP.H3.abf) %>%
  arrange(-PP3_4, -PP_4div3)
```
## check against brain mQTLs
```{r}
mQTLs_brain <- fread("~/mQTLs_all.txt.gz")

if (file.exists("pos_brain.txt.gz")) {
  pos_brain <- fread("pos_brain.txt.gz")
  setkey(pos_brain, "SNP")
} else {
  pos_brain <- unique(mQTLs_brain[, .(SNP = SNPid, CHR = paste0("chr", SNPchromosome), POS = SNPpos)])
  setkey(pos_brain, "SNP")
  fwrite(pos_brain, "pos_brain.txt.gz")
}

mQTLs_brain <- mQTLs_brain[, .(
  SNP = SNPid,
  gene = featureName,
  beta = SpearmanRho,
  `p-value` = pValue
)]
setkey(mQTLs_brain, "gene")

coloc_brain <- rbindlist(
  lapply(
    1:nrow(risk_scores),
    function(i) {
      localize_per_locus(
        i,
        mQTL_results = mQTLs_brain,
        snp_pos = pos_brain,
        N1 = 468,
        has_methy = FALSE
      )
    }
  )
)

coloc_brain_SNP <- rbindlist(
  lapply(
    1:nrow(risk_scores),
    function(i) {
      localize_per_locus(
        i,
        mQTL_results = mQTLs_brain,
        snp_pos = pos_brain,
        N1 = 468,
        has_methy = FALSE,
        return_summary = FALSE
      )
    }
  )
)
```
```{r}

coloc_brain %>%
  mutate(PP3_4 = PP.H3.abf + PP.H4.abf, PP_4div3 = PP.H4.abf / PP.H3.abf) %>%
  arrange(-PP3_4, -PP_4div3) %>%
  filter(PP3_4 > 0.99, PP_4div3 > 5)
coloc_brain %>%
  mutate(PP3_4 = PP.H3.abf + PP.H4.abf, PP_4div3 = PP.H4.abf / PP.H3.abf) %>%
  arrange(-PP3_4, -PP_4div3) %>%
  filter(PP3_4 > 0.99, PP_4div3 > 5) %>%
  distinct(probe, locus)
coloc_brain %>%
  mutate(PP3_4 = PP.H3.abf + PP.H4.abf, PP_4div3 = PP.H4.abf / PP.H3.abf) %>%
  arrange(-PP3_4, -PP_4div3) %>%
  filter(PP3_4 > 0.99, PP_4div3 > 5) %>%
  distinct(probe)
coloc_brain %>%
  mutate(PP3_4 = PP.H3.abf + PP.H4.abf, PP_4div3 = PP.H4.abf / PP.H3.abf) %>%
  arrange(-PP3_4, -PP_4div3) %>%
  filter(PP3_4 > 0.99, PP_4div3 > 5) %>%
  distinct(locus)
```
```{r}
coloc_brain %>%
  mutate(PP3_4 = PP.H3.abf + PP.H4.abf, PP_4div3 = PP.H4.abf / PP.H3.abf) %>%
  arrange(-PP3_4, -PP_4div3) %>%
  filter(PP3_4 > 0.99, PP_4div3 > 5) %>%
  write_delim("mqtl_pd_snp_colocalized_brain.txt")

coloc_brain %>%
  filter(PP.H4.abf > 0.9) %>%
  write_delim("mqtl_pd_snp_colocalized_brain_PH4.txt")

fwrite(coloc_brain_SNP, "mqtl_pd_snp_colocalization_brain.txt.gz", sep = "\t", row.names = F, quote = F)
```


## Visualizing colocalization

### Load data
```{r}
library(EnsDb.Hsapiens.v75)

res1_pval <- fread("mqtl_pd_snp_colocalized_blood.txt")
coloc_brain <- fread("mqtl_pd_snp_colocalized_brain.txt")
blood_cpg <- res1_pval %>%
  mutate(PP3_4 = PP.H3.abf + PP.H4.abf, PP_4div3 = PP.H4.abf / PP.H3.abf) %>%
  arrange(-PP3_4, -PP_4div3) %>%
  dplyr::filter(PP3_4 > 0.99, PP_4div3 > 5)
brain_cpg <- coloc_brain %>%
  mutate(PP3_4 = PP.H3.abf + PP.H4.abf, PP_4div3 = PP.H4.abf / PP.H3.abf) %>%
  arrange(-PP3_4, -PP_4div3) %>%
  dplyr::filter(PP3_4 > 0.99, PP_4div3 > 5)

top_blood <- as.character(blood_cpg[locus %in% c("69.0")]$probe)
top_brain <- as.character(brain_cpg[locus %in% c("69.0")]$probe)
top_coloc_loci <- unique(c(top_blood, top_brain))

top_2_blood <- as.character(blood_cpg[locus %in% c("70.0")]$probe)
top_2_brain <- as.character(brain_cpg[locus %in% c("70.0")]$probe)
top_2_coloc_loci <- unique(c(top_2_blood, top_2_brain))
probe_pos <- fread("terre_data/probe_pos.txt")
snp_pos <- fread("terre_data/snp_pos.txt", key = "SNP")
pos_brain <- fread("pos_brain.txt.gz")
setkey(pos_brain, "SNP")
sum_stats <- fread("~/nallsEtAl2019_excluding23andMe_allVariants.tab.gz", key = "SNP")

if (file.exists("mqtl_results_trunc.txt.gz")) {
  mQTL_results_trunc <- fread("mqtl_results_trunc.txt.gz")
  mQTLs_brain_trunc <- fread("mqtl_brain_trunc.txt.gz")
} else {
  fwrite(mQTL_results[gene %in% union(top_coloc_loci, top_2_coloc_loci)], "mqtl_results_trunc.txt.gz")
  fwrite(mQTLs_brain[gene %in% union(top_coloc_loci, top_2_coloc_loci)], "mqtl_brain_trunc.txt.gz")
}
edb <- EnsDb.Hsapiens.v75
unified_pos <- merge(pos_brain, snp_pos, by = c("CHR", "POS"), all = TRUE)
```
### Plot
```{r,fig.height=8.5, fig.height=11}

plot_cpgs_overlap <- function(cpgs) {
  # gather all SNPs as chr#:pos
  cpg_coord <- probe_pos[cpgs, on = "geneid"]
  cpg_snps <- unified_pos[CHR == cpg_coord$chr & POS > min(cpg_coord$s1) - 75e3 & POS < max(cpg_coord$s2) + 75e3]
  cgi <- UcscTrack(
    genome = "hg19", chromosome = unique(cpg_coord$chr), from = min(cpg_snps$POS),
    to = max(cpg_snps$POS), track = "cpgIslandExt", trackType = "AnnotationTrack",
    start = "chromStart", end = "chromEnd", id = "name", shape = "box", fill = "#006400",
    name = "CpG Islands"
  )
  gr <- GeneRegionTrack(
    getGeneRegionTrackForGviz(
      edb,
      featureIs = "gene_biotype",
      chromosome = gsub("chr", "", unique(cpg_coord$chr)),
      start = min(cpg_snps$POS), end = max(cpg_snps$POS),
    )
  )
  dat <- DataTrack(
    start = cpg_snps$POS,
    width = 1,
    data = rbind(
      -log10(mQTLs_brain_trunc[.(cpg_snps$SNP.x, cpgs), on = c("SNP", "gene")]$`p-value`),
      -log10(mQTL_results_trunc[.(cpg_snps$SNP.y, cpgs), on = c("SNP", "gene")]$`p-value`)
    ),
    chromosome = unique(cpg_coord$chr),
    genome = "hg19",
    groups = c("Brain", "Blood"),
    name = "-log10 mQTL P value"
  )
  gwas <- DataTrack(
    start = cpg_snps$POS,
    width = 1,
    data = -log10(sum_stats[paste0(cpg_snps$CHR, ":", cpg_snps$POS), on = "SNP"]$p),
    chromosome = unique(cpg_coord$chr),
    genome = "hg19",
    name = "-log10(GWAS P-value)"
  )
  plotTracks(
    list(gr, gwas, dat),
    collapseTranscripts = "meta", transcriptAnnotation = "symbol"
  )
}
# plot_cpgs_overlap(top_coloc_loci)
# plot_cpgs_overlap(top_2_coloc_loci)
plot_cpgs_overlap(union(top_coloc_loci, top_2_coloc_loci))
```
```{r}
library(ggbio)
library(biovizBase)
library(Homo.sapiens)
if (!"mQTLs_brain" %in% ls()) {
  mQTLs_brain <- fread("~/mQTLs_all.txt.gz")
  setkey(mQTLs_brain, "featureName")
}
if (!"mQTL_results" %in% ls()) {
  mQTL_results <- fread("terre_data/cis_all_impute_mQTL_results_9_methy_PC.txt.gz", key = "gene")
}

mQTL_results_pos <- mQTL_results[snp_pos, on = .(SNP)][probe_pos[, .(geneid, s1, s2)], on = .(gene = geneid)]

gr.brain_cpg <- transformDfToGr(na.omit(mQTLs_brain[top_blood, .(CHR = SNPchromosome, BP = SNPpos, P = pValue), on = "featureName"]), seqnames = "CHR", start = "BP", width = 1)
gr.brain_cpg$group <- "brain"
gr.blood_cpg <- transformDfToGr(na.omit(mQTL_results_pos[top_blood, .(CHR = as.numeric(gsub("chr", "", CHR)), BP = as.numeric(POS), P = `p-value`), on = "gene"]), seqnames = "CHR", start = "BP", width = 1)
gr.blood_cpg$group <- "blood"
chr_pattern <- "chr(.*):(.*)"
gwas_locus_1 <- sum_stats[, .(
  CHR = as.numeric(gsub(chr_pattern, "\\1", SNP)),
  BP = as.numeric(gsub(chr_pattern, "\\2", SNP)),
  P = p
)][CHR == gr.brain_cpg$CHR[1] & BP > gr.brain_cpg$BP[1] & BP < gr.brain_cpg$BP[length(gr.brain_cpg)]]
gr.pd_gwas <- transformDfToGr(gwas_locus_1, seqnames = "CHR", start = "BP", width = 1)
gr.pd_gwas$group <- "Parkinson's"
associations <- c(gr.blood_cpg, gr.brain_cpg, gr.pd_gwas)
seqlevelsStyle(associations) <- "UCSC"
genes <- subsetByOverlaps(genes(Homo.sapiens, columns = "SYMBOL"), associations)
p1 <- autoplot(c(gr.blood_cpg, gr.pd_gwas), geom = "point", coord = "genome", aes(color = group, y = -log10(P)))
p3 <- autoplot(gr.brain_cpg, geom = "point", coord = "genome", aes(y = -log10(P)))
p4 <- autoplot(gr.pd_gwas, geom = "point", coord = "genome", aes(y = -log10(P)))
p2 <- autoplot(Homo.sapiens, which = associations, columns = "SYMBOL", coord = "genome", stat = "identity")
tracks(p1, gene = p2)
```



#@TODO FIX ISSUE WITH BRAIN RANGE
